### 2020-12-08 Simulations.Rmd
update 1
first steps for season function, probably will change to a non game-by-game system soon

---
title: "Simulations"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
```

```{r}
# Read in and filter data
data <- read_csv("Data/game.csv")
data <- data %>% filter(data$season == 20132014, data$home_team_id %in% c(seq(1,30), 52), data$away_team_id %in% c(seq(1,30), 52), data$type == 'R')
data$home_team_id[data$home_team_id == 52] <- 11
data$away_team_id[data$away_team_id == 52] <- 11
```

# The Model
```{r}
# There are three key parameters: the probability that the worst team in the league beats the best team in a random game (given by 'p'), the standard deviation in game-to-game performance levels for each team (given by 'tao'), and the standard deviation in the team talent levels (given by 'sigma'). We as the researchers choose a value for p or the ratio of tao and sigma.
p = 0.25
tao = 4.28
sigma = 1

# Draw a "team quality" value for all teams.
teamQuality <- rnorm(30, 0, sigma)

# Calculate tie threshhold.
alpha <- qnorm(0.612, 0, sqrt(2 * (tao^2 + sigma^2)))
```

# Simulating a season
```{r}
simulateSeason <- function() {
  # Start all teams at 
  teamPoints <- rep(0, 30)
  for (game in 1:dim(data[1])) {
    homeTeam <- getHomeTeam(game)
    awayTeam <- getAwayTeam(game)
    # Extract game performance values
    homePerformance <- rnorm(teamQuality[homeTeam], tao)
    awayPerformance <- rnorm(teamQuality[awayTeam], tao)

    # Check if there is a tie
    tie <- abs(homePerformance - awayPerformance)

    # If the game did not go into overtime award points
    if (tie >= alpha) {
      if (homePerformance > awayPerformance) {
        teamPoints[homeTeam] = teamPoints[homeTeam] + 2
      }
      else {
        teamPoints[awayTeam] = teamPoints[awayTeam] + 2
      }
    }
    # If the game did go into overtime simulate outcome of overtime
    else {

    }

  }
}
```

# Helper function
```{r}
getHomeTeam <- function(index) {
  return (data$home_team_id[index])
}

getAwayTeam <- function(index) {
  return (data$away_team_id[index])
}
```

### 2020-12-08 Simulations.Rmd

@@
 simulateSeason <- function() {
-  # Start all teams at 
-  teamPoints <- rep(0, 30)
-  for (game in 1:dim(data[1])) {
+  # Create overall games matrix with format:
+  # GAME | TEAM 1 | TEAM 2 | GAMMA 1 | GAMMA 2 | GAMMA 1 - GAMMA 2 | OUTCOME 1 | OUTCOME 2
+  # Where 'outcome' is the number of points awarded to their respective teams
+  games <- list(seq(1, dim(data[1])[1]), rep(0, dim(data[1])[1]), rep(0, dim(data[1])[1]), rep(0, dim(data[1])[1]), rep(0, dim(data[1])[1]), rep(0, dim(data[1])[1]), rep(0, dim(data[1])[1]), rep(0, dim(data[1])[1]))
+
+  # Go through the games
+  for (game in 1:length(games[[1]])) {
+
+    # Get the teams for each game and put them in the games matrix
     homeTeam <- getHomeTeam(game)
     awayTeam <- getAwayTeam(game)
-    # Extract game performance values
-    homePerformance <- rnorm(teamQuality[homeTeam], tao)
-    awayPerformance <- rnorm(teamQuality[awayTeam], tao)
 
-    # Check if there is a tie
-    tie <- abs(homePerformance - awayPerformance)
+    games[[2]][game] <- homeTeam
+    games[[3]][game] <- awayTeam
 
-    # If the game did not go into overtime award points
-    if (tie >= alpha) {
-      if (homePerformance > awayPerformance) {
-        teamPoints[homeTeam] = teamPoints[homeTeam] + 2
+    # Extract game performance values -- insert into games matrix
+    homePerformance <- rnorm(n = 1, teamQuality[homeTeam], tao)
+    awayPerformance <- rnorm(n = 1, teamQuality[awayTeam], tao)
+
+    games[[4]][game] <- homePerformance
+    games[[5]][game] <- awayPerformance
+
+    # Check if there is a tie -- store difference for later
+    games[[6]][game] <- homePerformance - awayPerformance
+  }
+  a <- mean(games[[6]])
+  b <- max(games[[6]])
+
+  # Go through and calculate outcomes
+  for (game in 1:length(games[[1]])) {
+    # If there is a tie award 2 points to the 'winner'
+    if (abs(games[[6]][game]) >= alpha) {
+      if (games[[6]][game] >= 0) {
+        games[[7]][game] <- 2
       }
       else {
-        teamPoints[awayTeam] = teamPoints[awayTeam] + 2
+        games[[8]][game] <- 2
       }
     }
-    # If the game did go into overtime simulate outcome of overtime
+    # Else rescale performances and simulate a winner, giving two points to the winner and one to the loser
     else {
-
+      zeta <- (games[[6]][game] - a) / b
+      if (rbernoulli(1, p = zeta)) {
+        games[[7]][game] <- 2
+        games[[8]][game] <- 1
+      }
+      else {
+        games[[7]][game] <- 1
+        games[[8]][game] <- 2
+      }
+    }
   }
@@
 }
-}
 ```
 
-# Helper function
+# Helper functions
 ```{r}
@@
 }
+
+overtime <- function() {
+
+}
 ```

### 2020-12-09 Simulations.Rmd
Update Simulations.Rmd

@@
 
+# Read and filter the data
 ```{r}
-# Read in and filter data
 data <- read_csv("Data/game.csv")
@@
 alpha <- qnorm(0.612, 0, sqrt(2 * (tao^2 + sigma^2)))
+
 ```
@@
   }
-
+  return(games)
 }

### 2020-12-09 Simulations.Rmd
Update Simulations.Rmd

@@
 
+```{r}
+calcTotalPoints <- function(season){
+  totalPoints <- rep(0, 30)
+  for (game in 1:length(seasonGames[[1]])){
+    totalPoints[season[[2]][game]] = totalPoints[season[[2]][i]] +   season[[7]][game]
+    totalPoints[season[[3]][game]] = totalPoints[season[[3]][i]] + season[[8]][game]
+  }
+  return(totalPoints)
+}
+```
+
 # Helper functions

### 2020-12-09 Simulations.Rmd
Update Simulations.Rmd

--- 2020-12-09 Simulations.Rmd
@@
 
+```{r}
+calcTotalPoints <- function(season){
+  totalPoints <- rep(0, 30)
+  for (game in 1:length(seasonGames[[1]])){
+    totalPoints[season[[2]][game]] = totalPoints[season[[2]][i]] +   season[[7]][game]
+    totalPoints[season[[3]][game]] = totalPoints[season[[3]][i]] + season[[8]][game]
+  }
+  returnVec = list(totalPoints, c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild"))
+  return(returnVec)
+}
+```
+
+```{r}
+seperateConf(points){
+  eastern = rep(0, 16)
+  western = rep(0, 14)
+
+}
+```
+
 # Helper functions

### 2020-12-09 Simulations.Rmd
Update Simulations.Rmd
seperateConf

--- 2020-12-09 Simulations.Rmd
@@
 
+```{r}
+calcTotalPoints <- function(season){
+  totalPoints <- rep(0, 30)
+  for (game in 1:length(seasonGames[[1]])){
+    totalPoints[season[[2]][game]] = totalPoints[season[[2]][i]] +   season[[7]][game]
+    totalPoints[season[[3]][game]] = totalPoints[season[[3]][i]] + season[[8]][game]
+  }
+  returnVec = data.frame(totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild"))
+  return(returnVec)
+}
+```
+
+```{r}
+seperateConf <- function(points){
+  eastern = data.frame()
+  western = data.frame()
+  indexE = 1
+  indexW = 1
+  for (i in c(1:10, 12:15, 17, 29)){
+    eastern[indexE, 1] = points[i, 1]
+    eastern[indexE, 2] = points[i, 2]
+    indexE = indexE + 1
+  }
+
+  for (i in c(11, 16, 18:28, 30)){
+    western[indexW, 1] = points[i, 1]
+    western[indexW, 2] = points[i, 2]
+    indexW = indexW + 1
+  }
+  return(c(eastern, western))
+}
+```
+
 # Helper functions

### 2020-12-09 Simulations.Rmd

--- 2020-12-09 Simulations.Rmd
@@
 library(MASS)
+library(ggplot2)
 ```
@@
 
+nloop <- 1000
+oldGap <- rep(NA, nloop)
+newGap <- rep(NA, nloop)
+
+for (i in 1:nloop) {
+  season <- simulateSeason()
+  points <- calcTotalPoints(season)
+  oldEast <- returnOldEastern(points)
+  oldWest <- returnOldWestern(points)
+  newEast <- returnNewEastern(points)
+  newWest <- returnNewWestern(points)
+  oldGap[i] <- oldEast[8,2] - oldWest[8,2]
+  newGap[i] <- newEast[8,2] - newWest[8,2]
+}
+
+ggplot() +
+  geom_histogram(aes(oldGap) col = 2, bins = 50) +
+  geom_histogram(aes(newGap), col = 4, bins = 50)
+
+c(mean(oldGap), mean(newGap))
 ```
@@
 ```{r}
+# Get the sum of the points for each team. The return vector is sorted by team number value.
+calcTotalPoints <- function(season){
+  totalPoints <- rep(0, 30)
+  for (game in 1:length(season[[1]])){
+    totalPoints[season[[2]][game]] = totalPoints[season[[2]][game]] + season[[7]][game]
+    totalPoints[season[[3]][game]] = totalPoints[season[[3]][game]] + season[[8]][game]
+  }
+  returnVec = data.frame("id" = seq(1,30), totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild"))
+=======
 calcTotalPoints <- function(season){
@@
   }
-  return(totalPoints)
+  returnVec = data.frame(totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild"))
+
+  return(returnVec)
+}
+```
+
+```{r}
+
+returnOldEastern <- function(points) {
+  atlantic = points[points$id %in% c(1:5),]
+  northeast = points[points$id %in% c(6:10),]
+  southeast = points[points$id %in% c(11:15),]
+
+  atlantic = atlantic[order(-atlantic$totalPoints),]
+  northeast = northeast[order(-northeast$totalPoints),]
+  southeast = southeast[order(-southeast$totalPoints),]
+
+  wildcards = rbind(atlantic[3:5,],northeast[3:5,],southeast[3:5,])
+  wildcards = wildcards[order(-wildcards$totalPoints),]
+
+  return(rbind(atlantic[1:2,],northeast[1:2,],southeast[1:2,],wildcards[1:2,],wildcards[3:9,]))
+}
+
+returnOldWestern <- function(points) {
+  central = points[points$id %in% c(16:19, 29),]
+  northwest = points[points$id %in% c(20:23, 30),]
+  pacific = points[points$id %in% c(26:28),]
+
+  central = central[order(-central$totalPoints),]
+  northwest = northwest[order(-northwest$totalPoints),]
+  pacific = pacific[order(-pacific$totalPoints),]
+
+  wildcards = rbind(central[3:5,],northwest[3:5,],pacific[3:5,])
+  wildcards = wildcards[order(-wildcards$totalPoints),]
+
+  return(rbind(central[1:2,],northwest[1:2,],pacific[1:2,],wildcards[1:2,],wildcards[3:9,]))
+}
+
+returnNewEastern <- function(points) {
+  metro = points[points$id %in% c(1:5, 12, 15, 29),]
+  atlantic = points[points$id %in% c(6:10, 13, 14, 17),]
+
+  metro = metro[order(-metro$totalPoints),]
+  atlantic = atlantic[order(-atlantic$totalPoints),]
+
+  wildcards = rbind(metro[4:8,],atlantic[4:8,])
+  wildcards = wildcards[order(-wildcards$totalPoints),]
+
+  return(rbind(metro[1:3,],atlantic[1:3,],wildcards[1:2,],wildcards[3:10,]))
+}
+
+returnNewWestern <- function(points) {
+  central = points[points$id %in% c(11, 16, 18, 19, 21, 25, 30),]
+  pacific = points[points$id %in% c(20, 22, 23, 24, 26:28),]
+
+  central = central[order(-central$totalPoints),]
+  pacific = pacific[order(-pacific$totalPoints),]
+
+  wildcards = rbind(central[4:7,],pacific[4:7,])
+  wildcards = wildcards[order(-wildcards$totalPoints),]
+
+  return(rbind(central[1:3,],pacific[1:3,],wildcards[1:2,],wildcards[3:8,]))
+}
+```{r}
+seperateConf <- function(points){
+  eastern = data.frame()
+  western = data.frame()
+  indexE = 1
+  indexW = 1
+  for (i in c(1:10, 12:15, 17, 29)){
+    eastern[indexE, 1] = points[i, 1]
+    eastern[indexE, 2] = points[i, 2]
+    indexE = indexE + 1
+  }
+
+  for (i in c(11, 16, 18:28, 30)){
+    western[indexW, 1] = points[i, 1]
+    western[indexW, 2] = points[i, 2]
+    indexW = indexW + 1
+  }
+  return(c(eastern, western))
 }
@@
 }
-
-overtime <- function() {
-
-}
 ```

### 2020-12-09 Simulations.Rmd
Fix errors

@@
 library(ggplot2)
+set.seed(400)
 ```
@@
 
-nloop <- 1000
+nloop <- 10000
 oldGap <- rep(NA, nloop)
@@
 ggplot() +
-  geom_histogram(aes(oldGap) col = 2, bins = 50) +
+  geom_histogram(aes(oldGap), col = 2, bins = 50) +
   geom_histogram(aes(newGap), col = 4, bins = 50)
@@
   returnVec = data.frame("id" = seq(1,30), totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild"))
-=======
-calcTotalPoints <- function(season){
-  totalPoints <- rep(0, 30)
-  for (game in 1:length(seasonGames[[1]])){
-    totalPoints[season[[2]][game]] = totalPoints[season[[2]][i]] +   season[[7]][game]
-    totalPoints[season[[3]][game]] = totalPoints[season[[3]][i]] + season[[8]][game]
-  }
-  returnVec = data.frame(totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild"))
 
@@
 ```{r}
-
 returnOldEastern <- function(points) {
@@
 }
+```
+
 ```{r}

### 2020-12-09 Simulations.Rmd
Improve visualization

@@
 library(ggplot2)
-set.seed(400)
+library(plyr)
+set.seed(410)
 ```
@@
 
-nloop <- 10000
+nloop <- 1000
 oldGap <- rep(NA, nloop)
@@
 
-ggplot() +
-  geom_histogram(aes(oldGap), col = 2, bins = 50) +
-  geom_histogram(aes(newGap), col = 4, bins = 50)
+df <- data.frame(Rule_Type = rep(c("Old alignment & rules", "New alignment & rules"), each = nloop), gap = c(oldGap, newGap))
+dfDat <- ddply(df, "Rule_Type", summarise, grp.mean=mean(gap))
+
+ggplot(df, aes(x = gap, fill = Rule_Type)) +
+  geom_density(alpha = 0.4) +
+  geom_vline(aes(xintercept = mean(oldGap)), col = 4) +
+  geom_vline(data = dfDat, aes(xintercept = grp.mean, color = Rule_Type), linetype = "dashed") +
+  xlab("Conference gap") +
+  ylab("Density")
 

### 2020-12-09 slides.Rmd
Update slides.Rmd

---
title: "insert title here"
author:
output: powerpoint_presentation

---

### 2020-12-09 Simulations.Rmd
fix seeding formula
may still be wrong, trying to talk to people who know hockey more

@@
 library(plyr)
-set.seed(410)
+set.seed(400)
 ```
@@
 # The Model
-```{r}
+```{r, set.seed(0)}
 # There are three key parameters: the probability that the worst team in the league beats the best team in a random game (given by 'p'), the standard deviation in game-to-game performance levels for each team (given by 'tao'), and the standard deviation in the team talent levels (given by 'sigma').
@@
 
-  return(rbind(atlantic[1:2,],northeast[1:2,],southeast[1:2,],wildcards[1:2,],wildcards[3:9,]))
+  top8 <- rbind(atlantic[1:2,],northeast[1:2,],southeast[1:2,],wildcards[1:2,])
+  top8 = top8[order(-top8$totalPoints),]
+
+  return(rbind(top8, wildcards[3:9,]))
 }
@@
 
-  return(rbind(central[1:2,],northwest[1:2,],pacific[1:2,],wildcards[1:2,],wildcards[3:9,]))
+  top8 <- rbind(central[1:2,],northwest[1:2,],pacific[1:2,],wildcards[1:2,])
+  top8 = top8[order(-top8$totalPoints),]
+
+  return(rbind(top8, wildcards[3:9,]))
 }
@@
 
-  return(rbind(metro[1:3,],atlantic[1:3,],wildcards[1:2,],wildcards[3:10,]))
+  top8 <- rbind(metro[1:3,],atlantic[1:3,],wildcards[1:2,])
+  top8 = top8[order(-top8$totalPoints),]
+
+  return(rbind(top8, wildcards[3:10,]))
 }
@@
 
-  return(rbind(central[1:3,],pacific[1:3,],wildcards[1:2,],wildcards[3:8,]))
+  top8 <- rbind(central[1:3,],pacific[1:3,],wildcards[1:2,])
+  top8 = top8[order(-top8$totalPoints),]
+
+  return(rbind(top8, wildcards[3:8,]))
 }

### 2020-12-09 slides.Rmd

@@
 output: powerpoint_presentation
+---
 
----
+# NHL Background
+
+## The National Hockey League (NHL)
+
+-Each team plays 82 games per season
+-This allows for every team to play each team at least twice
+
+## The changes for the 2013-14 season
+
+- Winnipeg Jets move from the Eastern to the Western confrence
+- Columbus Blue jackets and Detroit Redwings move from the Wet to the East\
+
+## The reasoning for the sampling
+
+- Changes left the West with 2 divisions of 14 teams and the East with 2 divisions of 16 teams
+- 8 teams from both confrences wer garunteed to make the playoffs
+- The sampling is meant to test wheter or not it is more difficult to reach the playoffs as an eastern confrence team
+
+# Extension 2017-18 season
+
+- This season added the Vegas Golden Knights
+- Now the west has 1 divion of 8 teams and one division of 7 teams
+- This creates new probabillities for the required points to make it to the playoffs

### 2020-12-09 Simulations.Rmd
Fix algorithm

@@
 
-  top8 <- rbind(atlantic[1:2,],northeast[1:2,],southeast[1:2,],wildcards[1:2,])
-  top8 = top8[order(-top8$totalPoints),]
+  top6 <- rbind(atlantic[1:2,],northeast[1:2,],southeast[1:2,])
+  top6 = top6[order(-top6$totalPoints),]
 
-  return(rbind(top8, wildcards[3:9,]))
+  return(rbind(top6, wildcards))
 }
@@
 
-  top8 <- rbind(central[1:2,],northwest[1:2,],pacific[1:2,],wildcards[1:2,])
-  top8 = top8[order(-top8$totalPoints),]
+  top6 <- rbind(central[1:2,],northwest[1:2,],pacific[1:2,])
+  top6 = top6[order(-top6$totalPoints),]
 
-  return(rbind(top8, wildcards[3:9,]))
+  return(rbind(top6, wildcards))
 }
@@
 
-  top8 <- rbind(metro[1:3,],atlantic[1:3,],wildcards[1:2,])
-  top8 = top8[order(-top8$totalPoints),]
+  top6 <- rbind(metro[1:3,],atlantic[1:3,])
+  top6 = top6[order(-top6$totalPoints),]
 
-  return(rbind(top8, wildcards[3:10,]))
+  return(rbind(top6, wildcards))
 }
@@
 
-  top8 <- rbind(central[1:3,],pacific[1:3,],wildcards[1:2,])
-  top8 = top8[order(-top8$totalPoints),]
+  top6 <- rbind(central[1:3,],pacific[1:3,])
+  top6 = top6[order(-top$totalPoints),]
 
-  return(rbind(top8, wildcards[3:8,]))
+  return(rbind(top6, wildcards))
 }

### 2020-12-09 Simulations.Rmd
Update Simulations.Rmd

@@
 # The Model
-```{r, set.seed(0)}
+```{r, set.seed(400)}
 # There are three key parameters: the probability that the worst team in the league beats the best team in a random game (given by 'p'), the standard deviation in game-to-game performance levels for each team (given by 'tao'), and the standard deviation in the team talent levels (given by 'sigma').
@@
   top6 <- rbind(central[1:3,],pacific[1:3,])
-  top6 = top6[order(-top$totalPoints),]
+  top6 = top6[order(-top6$totalPoints),]
 

### 2020-12-09 Simulations.Rmd
fix
moved teamquality drawing to each season as it should be!

@@
 library(plyr)
-set.seed(400)
 ```
@@
 # The Model
-```{r, set.seed(400)}
+```{r}
+set.seed(400)
 # There are three key parameters: the probability that the worst team in the league beats the best team in a random game (given by 'p'), the standard deviation in game-to-game performance levels for each team (given by 'tao'), and the standard deviation in the team talent levels (given by 'sigma').
@@
 
-# Draw a "team quality" value for all teams.
-teamQuality <- rnorm(30, 0, sigma)
-
 # Calculate tie threshhold.
@@
 simulateSeason <- function() {
+  # Draw a "team quality" value for all teams.
+  teamQuality <- rnorm(30, 0, sigma)
+
   # Create overall games matrix with format:

### 2020-12-10 Simulations.Rmd
Added box & whisker plot

@@
 newGap <- rep(NA, nloop)
+seedPoints <- data.frame(integer())
 
@@
   newGap[i] <- newEast[8,2] - newWest[8,2]
+  allTeams <- rbind(newEast, newWest)
+  allTeams <- allTeams[order(-allTeams$totalPoints),]
+  seeds <- data.frame(seedPlace = seq(1,30), points = allTeams$totalPoints)
+  seedPoints <- rbind(seedPoints, seeds)
 }
@@
   geom_density(alpha = 0.4) +
-  geom_vline(aes(xintercept = mean(oldGap)), col = 4) +
   geom_vline(data = dfDat, aes(xintercept = grp.mean, color = Rule_Type), linetype = "dashed") +
@@
   ylab("Density")
+
+ggplot(seedPoints, aes(x = seedPlace, y = points, group = seedPlace)) +
+  geom_boxplot() +
+  coord_flip()
 

### 2020-12-10 slides.Rmd
Update slides.Rmd
Added to slides

--- 2020-12-09 slides.Rmd
@@
 author:
-output: powerpoint_presentation
+output: slidy_presentation 
+---
 
----
+## The National Hockey League (NHL)
+
+- Each team plays 82 games per season
+- This allows for every team to play each team at least twice
+- The league is split up into two conferences, the East and the West, each with two divisions
+-The East has the Metropolitan and the Atlantic divisions while the West has the Central and Pacific divisions.
+
+## The Changes for the 2013-14 season
+
+- Winnipeg Jets move from the Eastern to the Western confrence
+- Columbus Blue jackets and Detroit Redwings move from the West to the East
+- Each conference previoulsy had three divisions of five teams but these were condensed into the four divisions that currently exist
+
+# Playoff Structure
+- For the previous conference/division structure,
+
+# The Reasoning for the Sampling
+
+- Changes left the West with two divisions of 14 teams and the East with 2 divisions of 16 teams
+- 8 teams from both conferences were garunteed to make the playoffs
+- The sampling is meant to test whether or not it is more difficult to reach the playoffs as an eastern confrence team
+
+# Extension 2017-18 season
+
+- This season added the Vegas Golden Knights
+- Now the West has one divion of 8 teams and one division of 7 teams
+- This creates new probabilities for the required points to make it to the playoffs

### 2020-12-10 slides.tex
Added to slides

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={insert title here},
  pdfborder={0 0 0},
  breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\newif\ifbibliography
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-2}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{insert title here}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{NHL Background}
\protect\hypertarget{nhl-background}{}

\begin{block}{The National Hockey League (NHL)}

\begin{itemize}
\tightlist
\item
  Each team plays 82 games per season
\item
  This allows for every team to play each team at least twice
\end{itemize}

\end{block}

\begin{block}{The changes for the 2013-14 season}

\begin{itemize}
\tightlist
\item
  Winnipeg Jets move from the Eastern to the Western confrence
\item
  Columbus Blue jackets and Detroit Redwings move from the Wet to the
  East\\
\end{itemize}

\end{block}

\begin{block}{The reasoning for the sampling}

\begin{itemize}
\tightlist
\item
  Changes left the West with 2 divisions of 14 teams and the East with 2
  divisions of 16 teams
\item
  8 teams from both confrences wer garunteed to make the playoffs
\item
  The sampling is meant to test wheter or not it is more difficult to
  reach the playoffs as an eastern confrence team
\end{itemize}

\end{block}

\end{frame}

\begin{frame}{Extension 2017-18 season}
\protect\hypertarget{extension-2017-18-season}{}

\begin{itemize}
\tightlist
\item
  This season added the Vegas Golden Knights
\item
  Now the west has 1 divion of 8 teams and one division of 7 teams
\item
  This creates new probabillities for the required points to make it to
  the playoffs
\end{itemize}

\end{frame}

\end{document}

### 2020-12-10 slides.Rmd
Finished playoff structure slide

@@
 # Playoff Structure
-- For the previous conference/division structure,
+- For the previous conference/division structure, the top two teams in each division secure a berth into the playoffs
+- The remaining four slots are claimed by the two teams in each conference with the next highest point total regardles of the division. These are known as the wild cards.
+- The process is similar with the new divisions in each conference
+- Now the top three teams in each division secure a playoff berth and the wild cards are chosen using the same process as before
 

### 2020-12-10 Simulations.Rmd
remove useless files (new format)

@@
 ```{r}
-set.seed(400)
+set.seed(40)
 # There are three key parameters: the probability that the worst team in the league beats the best team in a random game (given by 'p'), the standard deviation in game-to-game performance levels for each team (given by 'tao'), and the standard deviation in the team talent levels (given by 'sigma').

### 2020-12-10 Simulations.Rmd
Update Simulations.Rmd
fix THE BUG

@@
 ```{r}
-set.seed(40)
+set.seed(405)
 # There are three key parameters: the probability that the worst team in the league beats the best team in a random game (given by 'p'), the standard deviation in game-to-game performance levels for each team (given by 'tao'), and the standard deviation in the team talent levels (given by 'sigma').
@@
 
-c(mean(oldGap), mean(newGap))
+c(mean(oldGap), mean(newGap), mean(newGap - oldGap))
 ```
@@
   northwest = points[points$id %in% c(20:23, 30),]
-  pacific = points[points$id %in% c(26:28),]
+  pacific = points[points$id %in% c(24:28),]
 

### 2020-12-10 slides.Rmd
remove useless files (new format)
fix THE BUG

@@
 
+# The Model: Goals
+
+- To simulate the performance of each team over a season with varying rules
+- To create an accurate scoring system that accounts for ties
+
+# The Model: Goals
+
+- To simulate the performance of each team over a season with varying rules
+- To create an accurate scoring system that accounts for ties
+- (For me) to learn how hockey works!
+
+# The Model: Methods
+
+- We generate each team's "quality" from a normal distribution
+- For each of the games played we generate a "game performance" metric from a normal distribution with the mean being the respective team's quality.
+
+![Quality Metric](Images/teamQuality.png)
+![Game Performance](Images/teamPerformance.png)
+
+# The Model: Methods
+
+- A game will go into overtime ~ 22.4% of the time
+- If the teams skills are within a threshold alpha, it will go into overtime
+- Once in overtime rescale performances and determine winner by weighted coinflip
+
+# The Model: Methods
+
+- We simulate 10,000 seasons to gather our metrics
+- Sort teams into the right conference, then into the right division based on which set of rules we are looking at (old or new)
+- Calculate the conference gap
+
+# 2013-2014 Season Results
+
+- We have found
+
 # Extension 2017-18 season

### 2020-12-10 slides.Rmd
added more slides

@@
 ---
-title: "insert title here"
-author:
+title: "Using Monte Carlo to estimate the effects of NHL rule changes"
+author: "Kilbourne Charrier, Michael Bowers, Clinton, Sam"
 output: slidy_presentation 
@@
 - The league is split up into two conferences, the East and the West, each with two divisions
--The East has the Metropolitan and the Atlantic divisions while the West has the Central and Pacific divisions.
+- The East has the Metropolitan and the Atlantic divisions while the West has the Central and Pacific divisions.
 
@@
 
-![Quality Metric](Images/teamQuality.png)
-![Game Performance](Images/teamPerformance.png)
+![](Images/teamQuality.png)
+![](Images/teamPerformance.png)
 
@@
 
-- We have found
+- We have found a noticably large conference gap, with the mean of the old alignment being about -0.32 and the mean of the new alignment being about 2.18.
+
+![](Images/GapVis.png)
+
+# 2013-2014 Season Results
+
+- Another interesting graph!
+
+![](Images/SeedPlace.png)
 

### 2020-12-10 slides.Rmd
Added points slide
added playoff image

@@
 title: "Using Monte Carlo to estimate the effects of NHL rule changes"
-author: "Kilbourne Charrier, Michael Bowers, Clinton, Sam"
+author: "Kilbourne Charrier, Michael Bauers, Clinton, Sam"
 output: slidy_presentation 
@@
 
+# Point System
+- Instead of basing ranking off of a team's win-loss record, the NHL tracks points for each team
+- 2 points for win (regulation or overtime)
+- 1 point for overtime loss
+- 0 points for regulation loss
+
 # Playoff Structure
@@
 - Now the top three teams in each division secure a playoff berth and the wild cards are chosen using the same process as before
+
+![](Images/nhl-playoff-bracket.jpg)
 

### 2020-12-10 AltSeasonSimulation.Rmd
Add files via upload
Simulation for 2017-2018 season, copy of simulation.rmd with some minor modifications

--- 2020-12-10 Simulations.Rmd
@@
----
 title: "Simulations"
@@
 data <- read_csv("Data/game.csv")
-data <- data %>% filter(data$season == 20132014, data$home_team_id %in% c(seq(1,30), 52), data$away_team_id %in% c(seq(1,30), 52), data$type == 'R')
+
+data <- data %>% filter(data$season == 20172018, data$home_team_id %in% c(seq(1,30), 52:54), data$away_team_id %in% c(seq(1,30), 52:54), data$type == 'R')
+
 data$home_team_id[data$home_team_id == 52] <- 11
 data$away_team_id[data$away_team_id == 52] <- 11
-```
-
-# The Model
-```{r}
-set.seed(400)
-# There are three key parameters: the probability that the worst team in the league beats the best team in a random game (given by 'p'), the standard deviation in game-to-game performance levels for each team (given by 'tao'), and the standard deviation in the team talent levels (given by 'sigma').
-We as the researchers choose a value for p or the ratio of tao and sigma.
-p = 0.25
-tao = 4.28
-sigma = 1
-
-# Calculate tie threshhold.
-alpha <- qnorm(0.612, 0, sqrt(2 * (tao^2 + sigma^2)))
-
-nloop <- 1000
-oldGap <- rep(NA, nloop)
-newGap <- rep(NA, nloop)
-seedPoints <- data.frame(integer())
-
-for (i in 1:nloop) {
-  season <- simulateSeason()
-  points <- calcTotalPoints(season)
-  oldEast <- returnOldEastern(points)
-  oldWest <- returnOldWestern(points)
-  newEast <- returnNewEastern(points)
-  newWest <- returnNewWestern(points)
-  oldGap[i] <- oldEast[8,2] - oldWest[8,2]
-  newGap[i] <- newEast[8,2] - newWest[8,2]
-  allTeams <- rbind(newEast, newWest)
-  allTeams <- allTeams[order(-allTeams$totalPoints),]
-  seeds <- data.frame(seedPlace = seq(1,30), points = allTeams$totalPoints)
-  seedPoints <- rbind(seedPoints, seeds)
-}
-
-df <- data.frame(Rule_Type = rep(c("Old alignment & rules", "New alignment & rules"), each = nloop), gap = c(oldGap, newGap))
-dfDat <- ddply(df, "Rule_Type", summarise, grp.mean=mean(gap))
-
-ggplot(df, aes(x = gap, fill = Rule_Type)) +
-  geom_density(alpha = 0.4) +
-  geom_vline(data = dfDat, aes(xintercept = grp.mean, color = Rule_Type), linetype = "dashed") +
-  xlab("Conference gap") +
-  ylab("Density")
-
-ggplot(seedPoints, aes(x = seedPlace, y = points, group = seedPlace)) +
-  geom_boxplot() +
-  coord_flip()
-
-c(mean(oldGap), mean(newGap), mean(newGap - oldGap))
+data$home_team_id[data$home_team_id == 53] <- 27
+data$away_team_id[data$away_team_id == 53] <- 27
+data$home_team_id[data$home_team_id == 54] <- 31
+data$away_team_id[data$away_team_id == 54] <- 31
 ```
@@
   # Draw a "team quality" value for all teams.
-  teamQuality <- rnorm(30, 0, sigma)
+  teamQuality <- rnorm(31, 0, sigma)
 
@@
 calcTotalPoints <- function(season){
-  totalPoints <- rep(0, 30)
+  totalPoints <- rep(0, 31)
   for (game in 1:length(season[[1]])){
@@
   }
-  returnVec = data.frame("id" = seq(1,30), totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild"))
+  returnVec = data.frame("id" = seq(1,31), totalPoints, "Team Name" = c("New Jersey Devils", "NY Islanders", "NY Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Boston Bruins", "Buffalo Sabres", "Montreal Canadiens", "Ottowa Senators", "Toronto Maple Leafs", "Winnipeg Jets", "Carolina Hurricanes", "Florida Panthers", "Tampa Bay Lightning", "Washington Capitals", "Chicago Blackhawks", "Detroit Red Wings", "Nashville Predators", "St Louis Blues", "Calgary Flames", "Colorado Avalanche", "Edmonton Oilers", "Vancouver Canucks", "Anaheim Ducks", "Dallas Stars", "LA Kings", "Phoenix Coyotes", "San Jose Sharks", "Columbus Blue Jackets", "Minnesota Wild","Vegas Golden Knights"))
 
@@
   central = points[points$id %in% c(11, 16, 18, 19, 21, 25, 30),]
-  pacific = points[points$id %in% c(20, 22, 23, 24, 26:28),]
+  pacific = points[points$id %in% c(20, 22, 23, 24, 26:28,31),]
 
@@
 
-  wildcards = rbind(central[4:7,],pacific[4:7,])
+  wildcards = rbind(central[4:7,],pacific[4:8,])
   wildcards = wildcards[order(-wildcards$totalPoints),]
@@
 
-  for (i in c(11, 16, 18:28, 30)){
+  for (i in c(11, 16, 18:28, 30:31)){
     western[indexW, 1] = points[i, 1]
@@
 ```
+
+# The Model
+```{r}
+set.seed(400)
+# There are three key parameters: the probability that the worst team in the league beats the best team in a random game (given by 'p'), the standard deviation in game-to-game performance levels for each team (given by 'tao'), and the standard deviation in the team talent levels (given by 'sigma').
+We as the researchers choose a value for p or the ratio of tao and sigma.
+p = 0.25
+tao = 4.28
+sigma = 1
+
+# Calculate tie threshhold.
+alpha <- qnorm(0.612, 0, sqrt(2 * (tao^2 + sigma^2)))
+
+nloop <- 1000
+oldGap <- rep(NA, nloop)
+newGap <- rep(NA, nloop)
+seedPoints <- data.frame(integer())
+
+for (i in 1:nloop) {
+  season <- simulateSeason()
+  points <- calcTotalPoints(season)
+  oldEast <- returnOldEastern(points)
+  oldWest <- returnOldWestern(points)
+  newEast <- returnNewEastern(points)
+  newWest <- returnNewWestern(points)
+  oldGap[i] <- oldEast[8,2] - oldWest[8,2]
+  newGap[i] <- newEast[8,2] - newWest[8,2]
+  allTeams <- rbind(newEast, newWest)
+  allTeams <- allTeams[order(-allTeams$totalPoints),]
+  seeds <- data.frame(seedPlace = seq(1,31), points = allTeams$totalPoints)
+  seedPoints <- rbind(seedPoints, seeds)
+}
+
+df <- data.frame(Rule_Type = rep(c("Old alignment & rules", "New alignment & rules"), each = nloop), gap = c(oldGap, newGap))
+dfDat <- ddply(df, "Rule_Type", summarise, grp.mean=mean(gap))
+
+ggplot(df, aes(x = gap, fill = Rule_Type)) +
+  geom_density(alpha = 0.4) +
+  geom_vline(data = dfDat, aes(xintercept = grp.mean, color = Rule_Type), linetype = "dashed") +
+  xlab("Conference gap") +
+  ylab("Density")
+
+ggplot(seedPoints, aes(x = seedPlace, y = points, group = seedPlace)) +
+  geom_boxplot() +
+  coord_flip()
+
+c(mean(oldGap), mean(newGap), mean(newGap - oldGap))
+```

### 2020-12-10 slides.Rmd
got rid of errors

@@
 title: "Using Monte Carlo to estimate the effects of NHL rule changes"
-author: "Kilbourne Charrier, Michael Bauers, Clinton, Sam"
+author: "Kilbourne, Michael, Clinton, Sam"
 output: slidy_presentation 
@@
 
-- We have found a noticably large conference gap, with the mean of the old alignment being about -0.32 and the mean of the new alignment being about 2.18.
+- We have found a noticably large conference gap, with the mean of the old alignment being about -0.0202 and the mean of the new alignment being about 2.30.
 
@@
 
+# What This Shows
+
+- Based on this model we can predict that it is more likely that a team in the West will make the playoffs with a worse overall score than another team in the East much more often than a team in the East will make the playoffs with a worse overall score than another team in the West.
+
+![](Images/GapVis.png)
+
 # Extension 2017-18 season
@@
 - This creates new probabilities for the required points to make it to the playoffs
+
+# Sources
+
+- Pettigrew, Stephen.
+“How the West Will Be Won: Using Monte Carlo Simulations to Estimate the Effects of NHL Realignment.” Journal of Quantitative Analysis in Sports, vol.
+10, no.
+3, 1 Jan.
+2014, 10.1515/jqas-2013-0125.
+Accessed 10 Dec.
+2020.

### 2020-12-10 slides.Rmd

@@
 
+# Conclusion
+
+- The rule changes put in place in 2013 did make it less "fair" for the East teams which have a harder time making the playoffs in comparison to the West teams.
+- Adding an additional team to the West shortened the conference gap by essentially increacing 'n'!
+- The more closely the two conferences resemble each other in terms of number of teams and number of districts the more "fair" the playoff system is.
+
 # Sources

### 2020-12-10 slides.Rmd
Add files via upload
added 2017-2018 graphics and results

--- 2020-12-10 slides.Rmd
@@
 title: "Using Monte Carlo to estimate the effects of NHL rule changes"
-author: "Kilbourne, Michael, Clinton, Sam"
+author: "Kilbourne Charrier, Michael Bauers, Clinton, Sam Downs"
 output: slidy_presentation 
@@
 
+# Results of 2017-2018
+
+- Conference gap under the old alignment was estimated to be 0.038 
+- Conference gap under the new alignment with 31 teams was estimated to be 1.091
+
+# 2017-2018 Plots
+
+![](Images/gapvis20172018.png)
+
+# 2017-2018 Plots
+
+![](Images/seedplace20172018.png)
+
 # Sources

### 2020-12-10 slides.Rmd
re add conclusion slide

@@
 
+# Conclusion
+
+- The rule changes put in place in 2013 did make it less "fair" for the East teams which have a harder time making the playoffs in comparison to the West teams.
+- Adding an additional team to the West shortened the conference gap by essentially increacing 'n'!
+- The more closely the two conferences resemble each other in terms of number of teams and number of districts the more "fair" the playoff system is.
+
+![](Images/GapVis.png)
+![](Images/gapvis20172018.png)
+
 # Sources

### 2020-12-10 slides.Rmd
fix slides
Added conference gap slide

@@
 - The sampling is meant to test whether or not it is more difficult to reach the playoffs as an eastern confrence team
+
+# Conference Gap Definition
+
+- After seeding the teams in each conference, we look at both of the 8th seeds. 
+- The 8th seed is the cutoff for which teams make the playoffs 
+- The difference between the point totals for the East 8th seed and the point totals for the West 8th seed is the conference gap.
 
@@
 ![](Images/GapVis.png)
+![](Images/SeedPlace.png)
 

### 2020-12-10 slides.Rmd
Update slides.Rmd

@@
 - The rule changes put in place in 2013 did make it less "fair" for the East teams which have a harder time making the playoffs in comparison to the West teams.
-- Adding an additional team to the West shortened the conference gap by essentially increacing 'n'!
+- Adding an additional team to the West shortened the conference gap
 - The more closely the two conferences resemble each other in terms of number of teams and number of districts the more "fair" the playoff system is.

### 2020-12-14 Paper.Rmd
get paper rmd created

---
title: "NHL realignment effects on conference gap"
author: "Kilbourne Charrier, Michael Bauers, Clinton, Sam Downs"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Motivation

## Introduction

# Methodology

# Results

### 2020-12-14 Paper.Rmd
Update Paper.Rmd

@@
 # Motivation
+  The endeavor to start this project began with the struggles given by our previous project idea.
+Because of this we were left stranded, searching fo a new project idea.
+When we learned that there was a paper written on the National Hockey League(NHL) with easy to access data, we hopped on the idea and got to work.
+As we continued our research and modeling we became more and more curious with the concept of wheighting confrences in sports by moving around teams.
+From the start, it already seemed unfair to have more teams in one confrence than the other.
+If every team had equal skill, the fourteen teams in the western confrence had an equal one four in seven chance of making it to the playoffs.
+In contrast, the sixteen teams in the eastern confrence had a fifty percent chance of making it to the playoffs.
+This uneven probabillity will be known for the rest of the paper as the confrence gap.
+  Now how could an organization as big as the NHL allow for that kind of discrepency between the confrences?
+That is what had us curious.
+As a continuation of our research we wanted to look at a nother season with a similiar balance shift in the confrences to see how that change couild have affected the outcome of the season.
+The season wwe choose was the recent expansion draft season of 2017-18.
+Before the start of the season, the NHL added a new franchis in the Las Vegas Golden Knights.
+This tewam was set to join the Western Confrence.
+This shifted the confrence split from fourteen to sixtten West to East to a closer fifteen to sixteen split.
+We wanted to see if our result was statistically significant from both the confrence gap of an even season and the confrence gap from our tested 2013-14 season.
 

### 2020-12-14 Paper.Rmd
write more, ~ 1 page now

--- 2020-12-14 Paper.Rmd
@@
 
+## Basic Idea
+
+The basic idea behind the model is that all of the teams have a "skill level" that they play at or near during any given game.
+Each game is competitive because we aren't sure how well either team will play, and even a team with a higher skill level can be beaten by another team for any number of reasons.
+We keep track of this skill level and draw other metrics based off of it, then use the metrics to compute a team's score for the season.
+This score is used to calculate the team's seed in their respective conference and division.
+
+## Creating the Model
+
+To start out, we first need to make sure that we calculate the game-to-game performance levels given by $\tau$ and the standard deviation of the team's talent levels given by $\sigma$.
+These are related to the following ratio, where p is the probability that the worst team in the league beats the best team in the league.
+
+$-\frac{2.8854}{\Phi^{-1}(p)} = \frac{\tau}{\sigma}$
+
+Based on the data from all of the seasons with 30 teams, we can calculate that the team with the worst record in the league beat the team with the best team about a quarter of the time.
+We choose to set p equal to 0.25 for this reason, making the ratio $\tau/\sigma$ = 4.28.
+This means that there are four times the variability in the team's performances in game than there is in the talent levels of the teams.
+The actual values of $\tau$ and $\sigma$ don't matter, just the ratio, but for simplicity we will set $\tau$ equal to 4.28 and $\sigma$ equal to 1.
+
+We then want to simulate a skill level metric for each of the thirty teams. We create the metric from the following distribution, with team i being equal to $\mu_{i}$.
+
+$\mu_{i}$~$N(0,\sigma^{2})$
+
+We draw the values from a normal distribution because in sports the overall performace of the teams is generally normally distributed.
+That is, there are a few really good teams, a few really bad teams, and a lot of teams in between.
+
+We hold the team skill levels constant throughout the season, this make intuitive sense, because there are very seldom large changes to the team's rosters or strategies throughout a season, but after each season teams often make at least some change.
+
+For each game that is played in the season we draw a performance metric for both competing teams.
+This metric is drawn from a normal distribution with the mean being equal to the team's skill level and the standard deviation being equal to $\tau^{2}$.
+This is based on the idea that a team with any skill level will play somewhere around their skill level, rarely playing much worse and rarely playing much better.
+This metric is calculated as such, with i indicating the home team and j indicating the away team.
+
+$\gamma_{i}$~$N(\mu_{i}, \tau^{2})$
+
+$\gamma_{j}$~$N(\mu_{j}, \tau^{2})$
+
 # Results

### 2020-12-14 Paper.Rmd
Update Paper.Rmd

--- 2020-12-14 Paper.Rmd
@@
 title: "NHL realignment effects on conference gap"
-author: "Kilbourne Charrier, Michael Bauers, Clinton, Sam Downs"
+author: "Kilbourne Charrier, Michael Bauers, Clinton Rothschild, Sam Downs"
 output: pdf_document
@@
 # Motivation
-  The endeavor to start this project began with the struggles given by our previous project idea.
-Because of this we were left stranded, searching fo a new project idea.
-When we learned that there was a paper written on the National Hockey League(NHL) with easy to access data, we hopped on the idea and got to work.
-As we continued our research and modeling we became more and more curious with the concept of wheighting confrences in sports by moving around teams.
+   In the 2013-14 National Hockey League(NHL) season, the league made some changes to the confrence numbers.
+The Winnipeg Jets moved from the Eastern confrence to the Western confrence.
+This createed an imbalance in the confrence numbers giving the Eastern confrence  teams and leaving the Western confrence with only fourteen teams.
 From the start, it already seemed unfair to have more teams in one confrence than the other.
@@
 ## Introduction
+  In 2011 when the Atlanta Thrashers became the Winnipeg Jets, a problem was presented as Winnipeg is considerably far from the east coast of the North american continent.
+Because of this and how the season is set up, the average travel time for teams playing winnipeg were fairly large as eastern confrence teams play echother more often.
+Just before the start of the 2013-14 season, the league offices for the NHL decided to move over the Winnipeg Jets from the Eastern Confrence to the Western Confrence and in return the Eastern Confrence aquired the Columbus Blue Jackets and the Detroit Redwings.
+This left the the two confrences at an uneven balance.
+The Eastern Confrence now had sixteen teams while the Western Confrence only had fourteen teams.
+This presented a curious problem as for both confrences, eight teams make the playoff brackets.
+It is easy to see the problem at hand that the western confrence has just as many playoff slots for having two less teams.
+The goal of our research and the paer we leaned on, was to discover what that difference could mean in the grander sense of the league moving forward with this set up.
+We did this based on the average points and wins necessary to make the playoffs for both confrences.
+  To further our study, we also looked at the 2017-18 season.
+This season is important becasue it gave the introduction of a new team in the league none as the Las Vegas Golden Knights.
+This team was added to the Western Confrence makiing the split slightly more even.
+Our goal for this was to see how adding a single team could change the skill outcome for the playoffs.
+We looked at the same measures to understand the new confrence gap as our previous analysis of the 2013-14 season.
 

### 2020-12-14 Paper.Rmd
finish? my part

--- 2020-12-14 Paper.Rmd
@@
 
+Logically, the winner of each game would be decided by the higher value of $\gamma$, but in hockey it is possible for the game to go into overtime.
+This happens in about 22.4% of the games that were played over the past 15 seasons, so we must solve the following equasion for $\alpha$:
+
+$Pr(|\gamma_{i}-\gamma_{j}| < \alpha) = 22.4$%
+
+Solving this equasion gives us a value of $\alpha \approx 1.769.$
+
+If there is a game in which $|\gamma_{i}-\gamma_{j}|\geq \alpha$, the team with the higher $\gamma$ value is awarded a win and is therefore given two points while the team with the lower $\gamma$ value is given 0 points.
+For games in which $|\gamma_{i}-\gamma_{j}| < \alpha$, the game is considered a tie.
+
+In the case of a tie we linearly resale the $\gamma$ values with the following formula:
+
+$\zeta = \frac{(\gamma_{i}-\gamma_{j}) - a}{b}$
+
+Where a is the mean and b is the maximum of all $\gamma_{i}-\gamma_{j}$ values from that season.
+
+Using the resulting value of $\zeta$ we use a Bernoulli distribution to simulate a weighted coin flip, with a success meaning that the home team wins and a failure meaning the away team wins.
+
+Winner ~ $Bernoulli(\zeta_{ij})$
+
+The winning team receives two points while the losing team receives one point.
+
+We simulate 10,000 seasons using these principles and simulate the old and new rulesets by going through the different division brackets depending on the rules.
+We then calculate the conference gap for both the old and the new ruleset.
+We also save the number of points earned by each seed to do analysis on later.
+
+## Additional model details
+
+Our model does not account for the real-world skill of teams, but because we are not doing any inference on specific team performance depending on rule changes this is not relevant.
+Because we are interested in the difference in scores of the 8th seeds we actually don't even need any real-world teams and can instead replace them with placeholders or just numbers.
+
+An alternative choice we could have made is that in the event of a tie we could have given either team a 50% chance of winning instead of linearly rescaling the values and using a weighted coin flip, but we don't know exactly how the outcome of NHL overtime games usually end up skewing -- whether it be in favor of the better team or if it is distributed more randomly.
+
 # Results
+
+# References
+
+- Pettigrew, Stephen.
+“How the West Will Be Won: Using Monte Carlo Simulations to Estimate the Effects of NHL Realignment.” Journal of Quantitative Analysis in Sports, vol.
+10, no.
+3, 1 Jan.
+2014, 10.1515/jqas-2013-0125.
+Accessed 10 Dec.
+2020.

### 2020-12-14 Paper.Rmd
upload my part
uploaded my part and fixed a lot of stuff

@@
 ---
-title: "NHL realignment effects on conference gap"
-author: "Kilbourne Charrier, Michael Bauers, Clinton, Sam Downs"
+title: "NHL realignment effects on the conference gap"
+author: "Kilbourne Charrier, Michael Bauers, Clinton Rothschild, Sam Downs"
 output: pdf_document
@@
 # Motivation
+   In the 2013-14 National Hockey League (NHL) season, the league made some changes to the conference numbers.
+The Winnipeg Jets moved from the Eastern conference to the Western conference.
+This created an imbalance in the conference numbers giving the Eastern conference sixteen teams and leaving the Western conference with only fourteen teams.
+From the start, it already seemed unfair to have more teams in one conference than the other.
+If every team had equal skill, the fourteen teams in the western conference had an equal one four in seven chance of making it to the playoffs.
+In contrast, the sixteen teams in the eastern conference had a fifty percent chance of making it to the playoffs.
+This uneven probabillity will be known for the rest of the paper as the conference gap.
+
+  Now how could an organization as big as the NHL allow for that kind of discrepency between the conferences?
+As a continuation of our research we wanted to look at another season with a similiar balance shift in the conferences to see how that change could have affected the outcome of the season.
+The season we chose was the recent expansion draft season of 2017-18.
+Before the start of the season, the NHL added a new franchise in the Las Vegas Golden Knights.
+This team was set to join the Western conference.
+This shifted the conference split from fourteen to sixtten West to East to a closer fifteen to sixteen split.
+We wanted to see if our result was statistically significant from both the conference gap of an even season and the conference gap from our tested 2013-14 season.
 
 ## Introduction
+  In 2011 when the Atlanta Thrashers became the Winnipeg Jets.
+A problem was presented as Winnipeg is considerably far from the east coast of the North American continent.
+Because of this and how the season is set up, the average travel time for teams playing winnipeg were fairly large as eastern conference teams play each other more often.
+Just before the start of the 2013-14 season, the league offices for the NHL decided to move over the Winnipeg Jets from the Eastern conference to the Western conference and in return the Eastern conference aquired the Columbus Blue Jackets and the Detroit Redwings.
+This left the the two conferences at an uneven balance.
+The Eastern conference now had sixteen teams while the Western conference only had fourteen teams.
+This presented a curious problem as for both conferences, eight teams make the playoff brackets.
+It is easy to see the problem at hand that the western conference has just as many playoff slots for having two less teams.
+The goal of our research and the paer we leaned on, was to discover what that difference could mean in the grander sense of the league moving forward with this set up.
+We did this based on the average points and wins necessary to make the playoffs for both conferences.
+
+  To further our study, we also looked at the 2017-18 season.
+This season is important becasue it gave the introduction of a new team in the league none as the Las Vegas Golden Knights.
+This team was added to the Western conference makiing the split slightly more even.
+Our goal for this was to see how adding a single team could change the skill outcome for the playoffs.
+We looked at the same measures to understand the new conference gap as our previous analysis of the 2013-14 season.
 

### 2020-12-14 Paper.Rmd

@@
 
+After doing 10,000 simulations of full NHL seasons using the Monte Carlo simulation methods, the results can be seen in the figure below.
+The red shaded graph represents the gap between the East and West 8th seeds (conference gap) under the conference and division alignment in the 2013-14 NHL season while the blue graph represents the gap between the East and West 8th seeds (conference gap) under the conference and division alignment in the 2012-2013 NHL season.
+The mean conference gap for the new alignment (2013-2014) is 2.30 points while the mean conference gap for the old alignment (2012-2013) is practically zero at -0.02 points.
+
+![](Images/GapVis.png)
+
+Looking at these results, the numbers fall in line with what would be the logical expectation under these circumstances.
+The new alignment moves the balance of teams over towards the East with 16 teams while the West sits at 14 teams.
+With more teams in the East, it makes sense that the East would have more teams with higher number of total points from the season.
+To look at these results in a different light, it is key to translate these numbers to the point scoring system of the NHL.
+On average, the difference between the conference gaps of the two division alignments comes out to be 2.32 in favor of the new alignment.
+The conference gap has been calculated as:
+
+East 8th Seed Total Points - West 8th Seed Total Points
+
+The mean conference gap, in terms of total points, means that the East’s 8th seeded team has on average 2.32 more points than the West’s 8th seeded team.
+In the NHL, a team earns 2 points for a win, 1 point for an overtime loss, and 0 points for a loss.
+With this point system, the East’s 8th seeded team on average needs to win one more game during the season or exchange two losses for two losses in overtime.
+This favors western conference teams in the way they have to earn less total points throughout the season to make the playoffs then eastern conference teams.
+
+Not only were the 8th seed from each conference calculated but each seed for the whole NHL.
+These box and whisker plots show a wide range of information for the total points for each seed.
+The endpoints of the whiskers represent the outlier point totals earned by each seed.
+The sides of the box represent the 2.5% and 97.5% quantiles.
+The line within the box represents the mean.
+The structure of this plot makes sense with the curve of less points the lower seed and more points the higher seed the team achieves.
+The higher number on the graph represents the lower seed of the team; 30 being the worst team in the league and 1 being the best team in the league.
+This has been cross-examined with real NHL data and the shape lines up in a similar fashion.
+
+![](Images/SeedPlace.png)
+
+# Conclusion
+
+Since the change to the dynamic of the NHL structure, there have been questions on which that structure causes an unfair advantage for one of the conferences: more specifically the Western Conference.
+With the West having two less teams, it was inferred that each of those teams would have a better chance of making the playoffs.
+They wouldn’t need to earn as many points to secure a playoff spot as teams in the Eastern Conference.
+After running 10,000 Monte Carlo simulations of the 2013-2014 NHL season, the results support the case that the Western Conference has a noticeable advantage over the Eastern Conference.
+The 8th seed in the east on average has an end of season point total of 2.32 points higher than the 8th seed in the west.
+This can cause problems because a statistically worse team can make the playoffs in the Western Conference while a better team might miss out in the Eastern Conference.
+This can be seen as unfair because the teams do not have any say in what conference they are in so this advantage is unwarranted.
+
 # References

### 2020-12-14 Paper.rmd
Add files via upload
Added parts to conclusion + methodology about 2017-2018 season, minor proof reading to tie things together

--- 2020-12-14 Paper.Rmd
@@
 ---
-title: "NHL realignment effects on the conference gap"
+title: "NHL Realignment Effects on Playoff Qualification"
 author: "Kilbourne Charrier, Michael Bauers, Clinton Rothschild, Sam Downs"
@@
 It is easy to see the problem at hand that the western conference has just as many playoff slots for having two less teams.
-The goal of our research and the paer we leaned on, was to discover what that difference could mean in the grander sense of the league moving forward with this set up.
+The goal of our analysis and the paper we based our ideas off of and partially replicated, was to discover what that difference could mean in the grander sense of the league moving forward with this set up.
 We did this based on the average points and wins necessary to make the playoffs for both conferences.
@@
   To further our study, we also looked at the 2017-18 season.
-This season is important becasue it gave the introduction of a new team in the league none as the Las Vegas Golden Knights.
-This team was added to the Western conference makiing the split slightly more even.
-Our goal for this was to see how adding a single team could change the skill outcome for the playoffs.
-We looked at the same measures to understand the new conference gap as our previous analysis of the 2013-14 season.
+This season is important because it gave the introduction of a new team in the league known as the Vegas Golden Knights.
+This team was introduced into the Pacific division of the Western conference, this meant that the balance between the number of teams in the East and West was more equal than the 2013-2014 split.
+Our goal for this was to investigate how this new conference structure could affect the conference gap.
+We used the same process to investigate this new season's structure as our previous analysis of the 2013-14 season.
+Given that the difference in the number of teams between conferences is less severe we suspected that our estimated conference gap under this setup (3 Divisions of 8 teams, 1 Division of 7 teams) would lie somewhere around the halfway point between the estimate for the pre 2013-2014 structure (6 Divisions of 5 teams each) and the 2013-2014 season (2 Divisions of 8 teams, 2 Divisions of 7 teams).
+
+## Pre 2013-2014 NHL Conference Structure
+
+![](Images/pre2013struct.png){width=60%}
+
+## 2013-2014 NHL Conference Structure
+
+![](Images/2013struct.png){width=60%}
+
+## 2017-2018 NHL Conference Structure
+
+![](Images/2017struct.png){width=60%}
 
@@
 
-The basic idea behind the model is that all of the teams have a "skill level" that they play at or near during any given game.
+The basic idea behind the model is that all of the teams have a base "skill level" that they play at or near during any given game throughout the season.
 Each game is competitive because we aren't sure how well either team will play, and even a team with a higher skill level can be beaten by another team for any number of reasons.
@@
 We hold the team skill levels constant throughout the season, this make intuitive sense, because there are very seldom large changes to the team's rosters or strategies throughout a season, but after each season teams often make at least some change.
+
+When drawing the season schedule for each simulation we chose just to use the same schedule as the real world NHL for that year.
+If we had taken into account some sort of measure of the performance of real world teams into our model this may have influenced the results to our research question.
+However since team skill is a random variable drawn at the start of each simulation with no relation to the real world skill of these teams this should not bias our results.
 
@@
 
-We simulate 10,000 seasons using these principles and simulate the old and new rulesets by going through the different division brackets depending on the rules.
-We then calculate the conference gap for both the old and the new ruleset.
+For each team the sum of points accumulated over the season was taken and teams were then sorted into conference and division.
+The qualifications for making the playoffs under each of our 3 seasons were then applied to the point totals for teams to determine seeding.
+Conference gap was then calculated by finding the difference in points between the 8th seed in the West and the 8th seed in the East.
+This process represents simulating a single season.
+
+We simulated 10,000 seasons for both the 2013-2014 and 2017-2018 analysis.
+We then calculate the conference gap for both the old and the new ruleset as the average conference gap across all simulations.
 We also save the number of points earned by each seed to do analysis on later.
@@
 Because we are interested in the difference in scores of the 8th seeds we actually don't even need any real-world teams and can instead replace them with placeholders or just numbers.
+The lack of a need for real world data was also an advantage to our analysis as it meant our results may be expressed as a general representation of the "fairness" of these conference structures since they are not related to the skill of real world teams.
 
@@
 
+When performing analysis on the 2017-2018 conference structure we chose to hold our parameters $p,\tau$ and $\sigma$ to be the same as stated above in the 2013-2014 analysis since we are talking about a reasonable small amount of time passing between these two seasons we thought it would be unreasonable to think that they would alter all that much.
+The same number of simulations were run also to make results more easily comparable.
+
 # Results
+
+After doing 10,000 simulations of full NHL seasons using the Monte Carlo simulation methods detailed above, the results can be seen in the figures below.
+The red shaded graph represents the gap between the East and West 8th seeds (conference gap) under the conference and division alignment in the 2013-14 NHL season while the blue graph represents the gap between the East and West 8th seeds (conference gap) under the conference and division alignment in the 2012-2013 NHL season.
+The mean conference gap for the new alignment (2013-2014) is 2.30 points while the mean conference gap for the old alignment (2012-2013) is practically zero at -0.02 points.
+
+![](Images/GapVis.png){width=85%}
+
+When looking at the 2017-2018 season we can see something close to what we hypothesized in our introduction, our conference gap does indeed lie somewhat close to halfway between the estimate for the 2013-2014 season and 2012-2013.
+When rerunning the analysis for 2017-2018 our estimate for conference gap under the 2012-2013 structure was 0.038 which is close to what was estimated above and makes sense as we expect the conferences to be equal in terms of likliehood of making the playoffs under these rules.
+The mean conference gap across our simulations for the 2017-2018 season was 1.091 which is approximately halfway between 0 and 2.3.
+It should be noticed that visually it can be seen that the density for the 2017-2018 season conference gap more closesly resembles the 2012-2013 density than the 2013-2014 density did.
+
+![](Images/gapvis20172018.png){width=85%}
+
+Looking at these results, the numbers fall in line with what would be the logical expectation under these circumstances.
+The 2013-2014 structure moves the balance of teams over towards the East with 16 teams while the West sits at 14 teams.
+With more teams in the East, it makes sense that the East would have more teams with higher number of total points from the season.
+To look at these results in a different light, it is key to translate these numbers to the point scoring system of the NHL.
+On average, the difference between the conference gaps of the two division alignments (2012-2013,2013-2014) comes out to be 2.32 in favor of the 2013-2014 alignment/structure.
+Under the 2017-2018 analyis our difference in conference gap between alignments comes out to be 1.053 (1.091-0.038).
+This result makes sense since the conferences are still unbalanced with the East having more teams than the West (16 and 15 respectively) but less unbalanced than 2013-2014 (16 and 14 respectively)  The conference gap has been calculated as:
+
+East 8th Seed Season Total Points - West 8th Seed Season Total Points
+
+The mean conference gap, in terms of total points, means that the East’s 8th seeded team has on average 2.30 more points than the West’s 8th seeded team for 2013-2014, similarly we expect the 8th seed in the Eastern conference to have 1.091 more points than their Western conference counterpart under the 2017-2018 structure.
+In the NHL, a team earns 2 points for a win, 1 point for an overtime loss, and 0 points for a loss.
+With this point system,under the 2013-2014 alignment the East’s 8th seeded team on average needs to win one more game during the season or exchange two losses for two losses in overtime.
+This favors Western conference teams in the way they have to earn less total points throughout the season to make the playoffs than Eastern conference teams.
+This same bias towards Western conference teams to make the playoffs is present under the 2017-2018 bias but the practical effects of a one point difference is not even worth a single win in an 82 game season, so although this effect is present it's not as drastic.
+
+Not only were the 8th seeds from each conference calculated but each seed for the whole NHL.
+These box and whisker plots show a wide range of information for the total points for each seed across the 10,000 simulated seasons for both the 2013-2014 and 2017-2018 alignemnts.
+The endpoints of the whiskers represent the outlier point totals earned by each seed.
+The sides of the box represent the 2.5% and 97.5% quantiles.
+The line within the box represents the mean.
+The structure of this plot makes sense with the curve of less points the lower seed and more points the higher seed the team achieves.
+The higher the number on the x axis the lower seed of the team; 30 being the worst team in the league (31 for 2017-2018 simulation) and 1 being the best team in the league.
+Interesting feautures of these plots are a tight spread of simulated points for teams seeded around the middle of the league.
+This makes sense given that these point values are probably closer to the mean point total for a season and thus more likely under the model we defined.
+Another note is that the overall range of point totals is greater in the 2013-2014 simulation than it is under the 2017-2018 simulation.
+This has been compared with real NHL data and the shape lines up in a similar fashion.
+
+## 2013-2014 Points by Seed
+![](Images/SeedPlace.png){width=80%}
+
+## 2017-2018 Points by Seed
+![](Images/seedplace20172018.png){width=80%}
+
+# Conclusion
+
+Since the change to the conference structure of the NHL to be unbalanced, there have been questions on whether these structures are giving an unfair advantage for one of the conferences: more specifically the Western Conference.
+With the West having two less teams under the 2013-2014 alignment, it was inferred that each of those teams would have a better chance of making the playoffs.
+They wouldn’t need to earn as many points to secure a playoff spot as teams in the Eastern Conference.
+After running 10,000 Monte Carlo simulations of the 2013-2014 NHL season, the results support the case that the Western Conference has a noticeable advantage over the Eastern Conference.
+The 8th seed in the east on average has an end of season point total of 2.3 (1.091 for 2017-2018) points higher than the 8th seed in the west.
+This can cause problems because a statistically worse team can make the playoffs in the Western Conference while a better team might miss out in the Eastern Conference.
+This is support of the idea that these rules introduce a bias agaisnt teams in the Eastern conference which goes agaisnt the league being "fair" to all teams.
 
@@
 2020.
+
+- History of Organizational Changes in the NHL. 26 Nov. 2020, en.wikipedia.org/wiki/History_of_organizational_changes_in_the_NHL. Accessed 10 Dec. 2020.

### 2020-12-14 Paper.Rmd
fix small errors

@@
 ## Introduction
-  In 2011 when the Atlanta Thrashers became the Winnipeg Jets.
+  In 2011 the Atlanta Thrashers became the Winnipeg Jets.
 A problem was presented as Winnipeg is considerably far from the east coast of the North American continent.
@@
 This presented a curious problem as for both conferences, eight teams make the playoff brackets.
-It is easy to see the problem at hand that the western conference has just as many playoff slots for having two less teams.
-The goal of our analysis and the paper we based our ideas off of and partially replicated, was to discover what that difference could mean in the grander sense of the league moving forward with this set up.
+It is easy to see the problem at hand that the western conference has just as many playoff slots for having two fewer teams.
+The goal of our analysis and the paper we based our ideas off of and partially replicated, was to discover what that difference could mean in the greater sense of the league moving forward with this set up.
 We did this based on the average points and wins necessary to make the playoffs for both conferences.

### 2022-02-15 Paper.Rmd
Fixed spelling errors

@@
 # Motivation
-   In the 2013-14 National Hockey League (NHL) season, the league made some changes to the conference numbers.
-The Winnipeg Jets moved from the Eastern conference to the Western conference.
-This created an imbalance in the conference numbers giving the Eastern conference sixteen teams and leaving the Western conference with only fourteen teams.
-From the start, it already seemed unfair to have more teams in one conference than the other.
-If every team had equal skill, the fourteen teams in the western conference had an equal one four in seven chance of making it to the playoffs.
-In contrast, the sixteen teams in the eastern conference had a fifty percent chance of making it to the playoffs.
-This uneven probabillity will be known for the rest of the paper as the conference gap.
+  Before the 2013-14 National Hockey League (NHL) season, the league made changes to the number of teams in each of the two conferences.
+The Winnipeg Jets moved from the Eastern Conference to the Western Conference which created an imbalance in the conference numbers giving the Eastern Conference sixteen teams and leaving the Western Conference with only fourteen teams.
+Since eight teams from each conference qualify for the playoffs, this imbalance created a “conference gap” that affected the probability of a team being selected for the playoffs.
 
-  Now how could an organization as big as the NHL allow for that kind of discrepency between the conferences?
-As a continuation of our research we wanted to look at another season with a similiar balance shift in the conferences to see how that change could have affected the outcome of the season.
+  How much would this change affect the outcome of the playoff selection process?
+This question was researched by Pettigrew1 in 2014 using a Monte Carlo approach and found the conference gap to be on average 2.74 points, meaning that teams in the Eastern Conference will have to win 1-2 games more than teams in the Western Conference in order to make the playoffs.
+For our research we validate Pettigrew’s paper as well as extend it by using a similar approach to compare this 2013-2014 season to another season with a similar balance shift in the conferences to see how the changes compare statistically.
+
   The season we chose was the recent expansion draft season of 2017-18.
-Before the start of the season, the NHL added a new franchise in the Las Vegas Golden Knights.
-This team was set to join the Western conference.
-This shifted the conference split from fourteen to sixtten West to East to a closer fifteen to sixteen split.
-We wanted to see if our result was statistically significant from both the conference gap of an even season and the conference gap from our tested 2013-14 season.
+Before the start of the season, the NHL added a new franchise in the Las Vegas Golden Knights which was set to join the Western conference.
+This added a team to the Western conference and resulted in a team split of fifteen to sixteen between East and West.
+Our goal was to see if the change in a team’s probability of making the playoffs was statistically significant when comparing between a year with even team distribution and the conference gap year from our tested 2013-14 and 2017-18  seasons.
 
 ## Introduction
-  In 2011 the Atlanta Thrashers became the Winnipeg Jets.
-A problem was presented as Winnipeg is considerably far from the east coast of the North American continent.
-Because of this and how the season is set up, the average travel time for teams playing winnipeg were fairly large as eastern conference teams play each other more often.
-Just before the start of the 2013-14 season, the league offices for the NHL decided to move over the Winnipeg Jets from the Eastern conference to the Western conference and in return the Eastern conference aquired the Columbus Blue Jackets and the Detroit Redwings.
-This left the the two conferences at an uneven balance.
-The Eastern conference now had sixteen teams while the Western conference only had fourteen teams.
-This presented a curious problem as for both conferences, eight teams make the playoff brackets.
-It is easy to see the problem at hand that the western conference has just as many playoff slots for having two fewer teams.
-The goal of our analysis and the paper we based our ideas off of and partially replicated, was to discover what that difference could mean in the greater sense of the league moving forward with this set up.
+  In 2011 the Atlanta Thrashers of the Eastern Conference became the Winnipeg Jets.
+This presented a problem as Winnipeg is a considerable distance from the east coast.
+Because of this and how the season game schedule is set up, the average travel time for teams playing Winnipeg from the Eastern Conference would have been fairly large.
+So, just before the start of the 2013-14 season, the league offices for the NHL decided to move the Winnipeg Jets from the Eastern Conference to the Western Conference and in return the Eastern Conference acquired the Columbus Blue Jackets and the Detroit Redwings.
+This left the two conferences with a different number of teams.
+The Eastern Conference then had sixteen teams while the Western Conference only had fourteen teams.
+This presented a curious problem since both conferences had eight teams advance to the playoffs at the end of the season.
+
+  The goal of our analysis and the paper we based our ideas off of and partially replicated1, was to discover how this impacted a team’s chances of making the playoffs.
 We did this based on the average points and wins necessary to make the playoffs for both conferences.
@@
 This season is important because it gave the introduction of a new team in the league known as the Vegas Golden Knights.
-This team was introduced into the Pacific division of the Western conference, this meant that the balance between the number of teams in the East and West was more equal than the 2013-2014 split.
+This team was introduced into the Pacific division of the Western conference, which meant that the balance between the number of teams in the East and West was more equal than the 2013-2014 split.
 Our goal for this was to investigate how this new conference structure could affect the conference gap.
-We used the same process to investigate this new season's structure as our previous analysis of the 2013-14 season.
+We used the same process to investigate this new season’s structure as our previous analysis of the 2013-14 season.
 Given that the difference in the number of teams between conferences is less severe we suspected that our estimated conference gap under this setup (3 Divisions of 8 teams, 1 Division of 7 teams) would lie somewhere around the halfway point between the estimate for the pre 2013-2014 structure (6 Divisions of 5 teams each) and the 2013-2014 season (2 Divisions of 8 teams, 2 Divisions of 7 teams).

### 2022-02-15 Paper.tex
Fixed spelling errors

--- 2020-12-14 Paper.rmd
@@
----
-title: "NHL Realignment Effects on Playoff Qualification"
-author: "Kilbourne Charrier, Michael Bauers, Clinton Rothschild, Sam Downs"
-output: pdf_document
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\title{NHL Realignment Effects on Playoff Qualification}
+\author{Kilbourne Charrier, Michael Bauers, Clinton Rothschild, Sam
+Downs}
+\date{}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-```
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={NHL Realignment Effects on Playoff Qualification},
+  pdfauthor={Kilbourne Charrier, Michael Bauers, Clinton Rothschild, Sam Downs},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 
-# Motivation
-   In the 2013-14 National Hockey League (NHL) season, the league made some changes to the conference numbers.
-The Winnipeg Jets moved from the Eastern conference to the Western conference.
-This created an imbalance in the conference numbers giving the Eastern conference sixteen teams and leaving the Western conference with only fourteen teams.
-From the start, it already seemed unfair to have more teams in one conference than the other.
-If every team had equal skill, the fourteen teams in the western conference had an equal one four in seven chance of making it to the playoffs.
-In contrast, the sixteen teams in the eastern conference had a fifty percent chance of making it to the playoffs.
-This uneven probabillity will be known for the rest of the paper as the conference gap.
+\begin{document}
+\maketitle
 
-  Now how could an organization as big as the NHL allow for that kind of discrepency between the conferences?
-As a continuation of our research we wanted to look at another season with a similiar balance shift in the conferences to see how that change could have affected the outcome of the season.
-The season we chose was the recent expansion draft season of 2017-18.
-Before the start of the season, the NHL added a new franchise in the Las Vegas Golden Knights.
-This team was set to join the Western conference.
-This shifted the conference split from fourteen to sixtten West to East to a closer fifteen to sixteen split.
-We wanted to see if our result was statistically significant from both the conference gap of an even season and the conference gap from our tested 2013-14 season.
+\hypertarget{motivation}{%
+\section{Motivation}\label{motivation}}
 
-## Introduction
-  In 2011 when the Atlanta Thrashers became the Winnipeg Jets.
-A problem was presented as Winnipeg is considerably far from the east coast of the North American continent.
-Because of this and how the season is set up, the average travel time for teams playing winnipeg were fairly large as eastern conference teams play each other more often.
-Just before the start of the 2013-14 season, the league offices for the NHL decided to move over the Winnipeg Jets from the Eastern conference to the Western conference and in return the Eastern conference aquired the Columbus Blue Jackets and the Detroit Redwings.
-This left the the two conferences at an uneven balance.
-The Eastern conference now had sixteen teams while the Western conference only had fourteen teams.
-This presented a curious problem as for both conferences, eight teams make the playoff brackets.
-It is easy to see the problem at hand that the western conference has just as many playoff slots for having two less teams.
-The goal of our analysis and the paper we based our ideas off of and partially replicated, was to discover what that difference could mean in the grander sense of the league moving forward with this set up.
-We did this based on the average points and wins necessary to make the playoffs for both conferences.
+In the 2013-14 National Hockey League (NHL) season, the league made some
+changes to the conference numbers. The Winnipeg Jets moved from the
+Eastern conference to the Western conference. This created an imbalance
+in the conference numbers giving the Eastern conference sixteen teams
+and leaving the Western conference with only fourteen teams. From the
+start, it already seemed unfair to have more teams in one conference
+than the other. If every team had equal skill, the fourteen teams in the
+western conference had an equal one four in seven chance of making it to
+the playoffs. In contrast, the sixteen teams in the eastern conference
+had a fifty percent chance of making it to the playoffs. This uneven
+probabillity will be known for the rest of the paper as the conference
+gap.
 
-  To further our study, we also looked at the 2017-18 season.
-This season is important because it gave the introduction of a new team in the league known as the Vegas Golden Knights.
-This team was introduced into the Pacific division of the Western conference, this meant that the balance between the number of teams in the East and West was more equal than the 2013-2014 split.
-Our goal for this was to investigate how this new conference structure could affect the conference gap.
-We used the same process to investigate this new season's structure as our previous analysis of the 2013-14 season.
-Given that the difference in the number of teams between conferences is less severe we suspected that our estimated conference gap under this setup (3 Divisions of 8 teams, 1 Division of 7 teams) would lie somewhere around the halfway point between the estimate for the pre 2013-2014 structure (6 Divisions of 5 teams each) and the 2013-2014 season (2 Divisions of 8 teams, 2 Divisions of 7 teams).
+Now how could an organization as big as the NHL allow for that kind of
+discrepency between the conferences? As a continuation of our research
+we wanted to look at another season with a similiar balance shift in the
+conferences to see how that change could have affected the outcome of
+the season. The season we chose was the recent expansion draft season of
+2017-18. Before the start of the season, the NHL added a new franchise
+in the Las Vegas Golden Knights. This team was set to join the Western
+conference. This shifted the conference split from fourteen to sixtten
+West to East to a closer fifteen to sixteen split. We wanted to see if
+our result was statistically significant from both the conference gap of
+an even season and the conference gap from our tested 2013-14 season.
 
-## Pre 2013-2014 NHL Conference Structure
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
-![](Images/pre2013struct.png){width=60%}
+In 2011 the Atlanta Thrashers became the Winnipeg Jets. A problem was
+presented as Winnipeg is considerably far from the east coast of the
+North American continent. Because of this and how the season is set up,
+the average travel time for teams playing winnipeg were fairly large as
+eastern conference teams play each other more often. Just before the
+start of the 2013-14 season, the league offices for the NHL decided to
+move over the Winnipeg Jets from the Eastern conference to the Western
+conference and in return the Eastern conference aquired the Columbus
+Blue Jackets and the Detroit Redwings. This left the the two conferences
+at an uneven balance. The Eastern conference now had sixteen teams while
+the Western conference only had fourteen teams. This presented a curious
+problem as for both conferences, eight teams make the playoff brackets.
+It is easy to see the problem at hand that the western conference has
+just as many playoff slots for having two fewer teams. The goal of our
+analysis and the paper we based our ideas off of and partially
+replicated, was to discover what that difference could mean in the
+greater sense of the league moving forward with this set up. We did this
+based on the average points and wins necessary to make the playoffs for
+both conferences.
 
-## 2013-2014 NHL Conference Structure
+To further our study, we also looked at the 2017-18 season. This season
+is important because it gave the introduction of a new team in the
+league known as the Vegas Golden Knights. This team was introduced into
+the Pacific division of the Western conference, this meant that the
+balance between the number of teams in the East and West was more equal
+than the 2013-2014 split. Our goal for this was to investigate how this
+new conference structure could affect the conference gap. We used the
+same process to investigate this new season's structure as our previous
+analysis of the 2013-14 season. Given that the difference in the number
+of teams between conferences is less severe we suspected that our
+estimated conference gap under this setup (3 Divisions of 8 teams, 1
+Division of 7 teams) would lie somewhere around the halfway point
+between the estimate for the pre 2013-2014 structure (6 Divisions of 5
+teams each) and the 2013-2014 season (2 Divisions of 8 teams, 2
+Divisions of 7 teams).
 
-![](Images/2013struct.png){width=60%}
+\hypertarget{pre-2013-2014-nhl-conference-structure}{%
+\subsection{Pre 2013-2014 NHL Conference
+Structure}\label{pre-2013-2014-nhl-conference-structure}}
 
-## 2017-2018 NHL Conference Structure
+\includegraphics[width=0.6\textwidth,height=\textheight]{Images/pre2013struct.png}
 
-![](Images/2017struct.png){width=60%}
+\hypertarget{nhl-conference-structure}{%
+\subsection{2013-2014 NHL Conference
+Structure}\label{nhl-conference-structure}}
 
-# Methodology
+\includegraphics[width=0.6\textwidth,height=\textheight]{Images/2013struct.png}
 
-## Basic Idea
+\hypertarget{nhl-conference-structure-1}{%
+\subsection{2017-2018 NHL Conference
+Structure}\label{nhl-conference-structure-1}}
 
-The basic idea behind the model is that all of the teams have a base "skill level" that they play at or near during any given game throughout the season.
-Each game is competitive because we aren't sure how well either team will play, and even a team with a higher skill level can be beaten by another team for any number of reasons.
-We keep track of this skill level and draw other metrics based off of it, then use the metrics to compute a team's score for the season.
-This score is used to calculate the team's seed in their respective conference and division.
+\includegraphics[width=0.6\textwidth,height=\textheight]{Images/2017struct.png}
 
-## Creating the Model
+\hypertarget{methodology}{%
+\section{Methodology}\label{methodology}}
 
-To start out, we first need to make sure that we calculate the game-to-game performance levels given by $\tau$ and the standard deviation of the team's talent levels given by $\sigma$.
-These are related to the following ratio, where p is the probability that the worst team in the league beats the best team in the league.
+\hypertarget{basic-idea}{%
+\subsection{Basic Idea}\label{basic-idea}}
 
-$-\frac{2.8854}{\Phi^{-1}(p)} = \frac{\tau}{\sigma}$
+The basic idea behind the model is that all of the teams have a base
+``skill level'' that they play at or near during any given game
+throughout the season. Each game is competitive because we aren't sure
+how well either team will play, and even a team with a higher skill
+level can be beaten by another team for any number of reasons. We keep
+track of this skill level and draw other metrics based off of it, then
+use the metrics to compute a team's score for the season. This score is
+used to calculate the team's seed in their respective conference and
+division.
 
-Based on the data from all of the seasons with 30 teams, we can calculate that the team with the worst record in the league beat the team with the best team about a quarter of the time.
-We choose to set p equal to 0.25 for this reason, making the ratio $\tau/\sigma$ = 4.28.
-This means that there are four times the variability in the team's performances in game than there is in the talent levels of the teams.
-The actual values of $\tau$ and $\sigma$ don't matter, just the ratio, but for simplicity we will set $\tau$ equal to 4.28 and $\sigma$ equal to 1.
+\hypertarget{creating-the-model}{%
+\subsection{Creating the Model}\label{creating-the-model}}
 
-We then want to simulate a skill level metric for each of the thirty teams. We create the metric from the following distribution, with team i being equal to $\mu_{i}$.
+To start out, we first need to make sure that we calculate the
+game-to-game performance levels given by \(\tau\) and the standard
+deviation of the team's talent levels given by \(\sigma\). These are
+related to the following ratio, where p is the probability that the
+worst team in the league beats the best team in the league.
 
-$\mu_{i}$~$N(0,\sigma^{2})$
+\(-\frac{2.8854}{\Phi^{-1}(p)} = \frac{\tau}{\sigma}\)
 
-We draw the values from a normal distribution because in sports the overall performace of the teams is generally normally distributed.
-That is, there are a few really good teams, a few really bad teams, and a lot of teams in between.
+Based on the data from all of the seasons with 30 teams, we can
+calculate that the team with the worst record in the league beat the
+team with the best team about a quarter of the time. We choose to set p
+equal to 0.25 for this reason, making the ratio \(\tau/\sigma\) = 4.28.
+This means that there are four times the variability in the team's
+performances in game than there is in the talent levels of the teams.
+The actual values of \(\tau\) and \(\sigma\) don't matter, just the
+ratio, but for simplicity we will set \(\tau\) equal to 4.28 and
+\(\sigma\) equal to 1.
 
-We hold the team skill levels constant throughout the season, this make intuitive sense, because there are very seldom large changes to the team's rosters or strategies throughout a season, but after each season teams often make at least some change.
+We then want to simulate a skill level metric for each of the thirty
+teams. We create the metric from the following distribution, with team i
+being equal to \(\mu_{i}\).
 
-When drawing the season schedule for each simulation we chose just to use the same schedule as the real world NHL for that year.
-If we had taken into account some sort of measure of the performance of real world teams into our model this may have influenced the results to our research question.
-However since team skill is a random variable drawn at the start of each simulation with no relation to the real world skill of these teams this should not bias our results.
+\(\mu_{i}\)\textasciitilde{}\(N(0,\sigma^{2})\)
 
-For each game that is played in the season we draw a performance metric for both competing teams.
-This metric is drawn from a normal distribution with the mean being equal to the team's skill level and the standard deviation being equal to $\tau^{2}$.
-This is based on the idea that a team with any skill level will play somewhere around their skill level, rarely playing much worse and rarely playing much better.
-This metric is calculated as such, with i indicating the home team and j indicating the away team.
+We draw the values from a normal distribution because in sports the
+overall performace of the teams is generally normally distributed. That
+is, there are a few really good teams, a few really bad teams, and a lot
+of teams in between.
 
-$\gamma_{i}$~$N(\mu_{i}, \tau^{2})$
+We hold the team skill levels constant throughout the season, this make
+intuitive sense, because there are very seldom large changes to the
+team's rosters or strategies throughout a season, but after each season
+teams often make at least some change.
 
-$\gamma_{j}$~$N(\mu_{j}, \tau^{2})$
+When drawing the season schedule for each simulation we chose just to
+use the same schedule as the real world NHL for that year. If we had
+taken into account some sort of measure of the performance of real world
+teams into our model this may have influenced the results to our
+research question. However since team skill is a random variable drawn
+at the start of each simulation with no relation to the real world skill
+of these teams this should not bias our results.
 
-Logically, the winner of each game would be decided by the higher value of $\gamma$, but in hockey it is possible for the game to go into overtime.
-This happens in about 22.4% of the games that were played over the past 15 seasons, so we must solve the following equasion for $\alpha$:
+For each game that is played in the season we draw a performance metric
+for both competing teams. This metric is drawn from a normal
+distribution with the mean being equal to the team's skill level and the
+standard deviation being equal to \(\tau^{2}\). This is based on the
+idea that a team with any skill level will play somewhere around their
+skill level, rarely playing much worse and rarely playing much better.
+This metric is calculated as such, with i indicating the home team and j
+indicating the away team.
 
-$Pr(|\gamma_{i}-\gamma_{j}| < \alpha) = 22.4$%
+\(\gamma_{i}\)\textasciitilde{}\(N(\mu_{i}, \tau^{2})\)
 
-Solving this equasion gives us a value of $\alpha \approx 1.769.$
+\(\gamma_{j}\)\textasciitilde{}\(N(\mu_{j}, \tau^{2})\)
 
-If there is a game in which $|\gamma_{i}-\gamma_{j}|\geq \alpha$, the team with the higher $\gamma$ value is awarded a win and is therefore given two points while the team with the lower $\gamma$ value is given 0 points.
-For games in which $|\gamma_{i}-\gamma_{j}| < \alpha$, the game is considered a tie.
+Logically, the winner of each game would be decided by the higher value
+of \(\gamma\), but in hockey it is possible for the game to go into
+overtime. This happens in about 22.4\% of the games that were played
+over the past 15 seasons, so we must solve the following equasion for
+\(\alpha\):
 
-In the case of a tie we linearly resale the $\gamma$ values with the following formula:
+\(Pr(|\gamma_{i}-\gamma_{j}| < \alpha) = 22.4\)\%
 
-$\zeta = \frac{(\gamma_{i}-\gamma_{j}) - a}{b}$
+Solving this equasion gives us a value of \(\alpha \approx 1.769.\)
 
-Where a is the mean and b is the maximum of all $\gamma_{i}-\gamma_{j}$ values from that season.
+If there is a game in which \(|\gamma_{i}-\gamma_{j}|\geq \alpha\), the
+team with the higher \(\gamma\) value is awarded a win and is therefore
+given two points while the team with the lower \(\gamma\) value is given
+0 points. For games in which \(|\gamma_{i}-\gamma_{j}| < \alpha\), the
+game is considered a tie.
 
-Using the resulting value of $\zeta$ we use a Bernoulli distribution to simulate a weighted coin flip, with a success meaning that the home team wins and a failure meaning the away team wins.
+In the case of a tie we linearly resale the \(\gamma\) values with the
+following formula:
 
-Winner ~ $Bernoulli(\zeta_{ij})$
+\(\zeta = \frac{(\gamma_{i}-\gamma_{j}) - a}{b}\)
 
-The winning team receives two points while the losing team receives one point.
+Where a is the mean and b is the maximum of all
+\(\gamma_{i}-\gamma_{j}\) values from that season.
 
-For each team the sum of points accumulated over the season was taken and teams were then sorted into conference and division.
-The qualifications for making the playoffs under each of our 3 seasons were then applied to the point totals for teams to determine seeding.
-Conference gap was then calculated by finding the difference in points between the 8th seed in the West and the 8th seed in the East.
-This process represents simulating a single season.
+Using the resulting value of \(\zeta\) we use a Bernoulli distribution
+to simulate a weighted coin flip, with a success meaning that the home
+team wins and a failure meaning the away team wins.
 
-We simulated 10,000 seasons for both the 2013-2014 and 2017-2018 analysis.
-We then calculate the conference gap for both the old and the new ruleset as the average conference gap across all simulations.
-We also save the number of points earned by each seed to do analysis on later.
+Winner \textasciitilde{} \(Bernoulli(\zeta_{ij})\)
 
-## Additional model details
+The winning team receives two points while the losing team receives one
+point.
 
-Our model does not account for the real-world skill of teams, but because we are not doing any inference on specific team performance depending on rule changes this is not relevant.
-Because we are interested in the difference in scores of the 8th seeds we actually don't even need any real-world teams and can instead replace them with placeholders or just numbers.
-The lack of a need for real world data was also an advantage to our analysis as it meant our results may be expressed as a general representation of the "fairness" of these conference structures since they are not related to the skill of real world teams.
+For each team the sum of points accumulated over the season was taken
+and teams were then sorted into conference and division. The
+qualifications for making the playoffs under each of our 3 seasons were
+then applied to the point totals for teams to determine seeding.
+Conference gap was then calculated by finding the difference in points
+between the 8th seed in the West and the 8th seed in the East. This
+process represents simulating a single season.
 
-An alternative choice we could have made is that in the event of a tie we could have given either team a 50% chance of winning instead of linearly rescaling the values and using a weighted coin flip, but we don't know exactly how the outcome of NHL overtime games usually end up skewing -- whether it be in favor of the better team or if it is distributed more randomly.
+We simulated 10,000 seasons for both the 2013-2014 and 2017-2018
+analysis. We then calculate the conference gap for both the old and the
+new ruleset as the average conference gap across all simulations. We
+also save the number of points earned by each seed to do analysis on
+later.
 
-When performing analysis on the 2017-2018 conference structure we chose to hold our parameters $p,\tau$ and $\sigma$ to be the same as stated above in the 2013-2014 analysis since we are talking about a reasonable small amount of time passing between these two seasons we thought it would be unreasonable to think that they would alter all that much.
-The same number of simulations were run also to make results more easily comparable.
+\hypertarget{additional-model-details}{%
+\subsection{Additional model details}\label{additional-model-details}}
 
-# Results
+Our model does not account for the real-world skill of teams, but
+because we are not doing any inference on specific team performance
+depending on rule changes this is not relevant. Because we are
+interested in the difference in scores of the 8th seeds we actually
+don't even need any real-world teams and can instead replace them with
+placeholders or just numbers. The lack of a need for real world data was
+also an advantage to our analysis as it meant our results may be
+expressed as a general representation of the ``fairness'' of these
+conference structures since they are not related to the skill of real
+world teams.
 
-After doing 10,000 simulations of full NHL seasons using the Monte Carlo simulation methods detailed above, the results can be seen in the figures below.
-The red shaded graph represents the gap between the East and West 8th seeds (conference gap) under the conference and division alignment in the 2013-14 NHL season while the blue graph represents the gap between the East and West 8th seeds (conference gap) under the conference and division alignment in the 2012-2013 NHL season.
-The mean conference gap for the new alignment (2013-2014) is 2.30 points while the mean conference gap for the old alignment (2012-2013) is practically zero at -0.02 points.
+An alternative choice we could have made is that in the event of a tie
+we could have given either team a 50\% chance of winning instead of
+linearly rescaling the values and using a weighted coin flip, but we
+don't know exactly how the outcome of NHL overtime games usually end up
+skewing -- whether it be in favor of the better team or if it is
+distributed more randomly.
 
-![](Images/GapVis.png){width=85%}
+When performing analysis on the 2017-2018 conference structure we chose
+to hold our parameters \(p,\tau\) and \(\sigma\) to be the same as
+stated above in the 2013-2014 analysis since we are talking about a
+reasonable small amount of time passing between these two seasons we
+thought it would be unreasonable to think that they would alter all that
+much. The same number of simulations were run also to make results more
+easily comparable.
 
-When looking at the 2017-2018 season we can see something close to what we hypothesized in our introduction, our conference gap does indeed lie somewhat close to halfway between the estimate for the 2013-2014 season and 2012-2013.
-When rerunning the analysis for 2017-2018 our estimate for conference gap under the 2012-2013 structure was 0.038 which is close to what was estimated above and makes sense as we expect the conferences to be equal in terms of likliehood of making the playoffs under these rules.
-The mean conference gap across our simulations for the 2017-2018 season was 1.091 which is approximately halfway between 0 and 2.3.
-It should be noticed that visually it can be seen that the density for the 2017-2018 season conference gap more closesly resembles the 2012-2013 density than the 2013-2014 density did.
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-![](Images/gapvis20172018.png){width=85%}
+After doing 10,000 simulations of full NHL seasons using the Monte Carlo
+simulation methods detailed above, the results can be seen in the
+figures below. The red shaded graph represents the gap between the East
+and West 8th seeds (conference gap) under the conference and division
+alignment in the 2013-14 NHL season while the blue graph represents the
+gap between the East and West 8th seeds (conference gap) under the
+conference and division alignment in the 2012-2013 NHL season. The mean
+conference gap for the new alignment (2013-2014) is 2.30 points while
+the mean conference gap for the old alignment (2012-2013) is practically
+zero at -0.02 points.
 
-Looking at these results, the numbers fall in line with what would be the logical expectation under these circumstances.
-The 2013-2014 structure moves the balance of teams over towards the East with 16 teams while the West sits at 14 teams.
-With more teams in the East, it makes sense that the East would have more teams with higher number of total points from the season.
-To look at these results in a different light, it is key to translate these numbers to the point scoring system of the NHL.
-On average, the difference between the conference gaps of the two division alignments (2012-2013,2013-2014) comes out to be 2.32 in favor of the 2013-2014 alignment/structure.
-Under the 2017-2018 analyis our difference in conference gap between alignments comes out to be 1.053 (1.091-0.038).
-This result makes sense since the conferences are still unbalanced with the East having more teams than the West (16 and 15 respectively) but less unbalanced than 2013-2014 (16 and 14 respectively)  The conference gap has been calculated as:
+\includegraphics[width=0.85\textwidth,height=\textheight]{Images/GapVis.png}
+
+When looking at the 2017-2018 season we can see something close to what
+we hypothesized in our introduction, our conference gap does indeed lie
+somewhat close to halfway between the estimate for the 2013-2014 season
+and 2012-2013. When rerunning the analysis for 2017-2018 our estimate
+for conference gap under the 2012-2013 structure was 0.038 which is
+close to what was estimated above and makes sense as we expect the
+conferences to be equal in terms of likliehood of making the playoffs
+under these rules. The mean conference gap across our simulations for
+the 2017-2018 season was 1.091 which is approximately halfway between 0
+and 2.3. It should be noticed that visually it can be seen that the
+density for the 2017-2018 season conference gap more closesly resembles
+the 2012-2013 density than the 2013-2014 density did.
+
+\includegraphics[width=0.85\textwidth,height=\textheight]{Images/gapvis20172018.png}
+
+Looking at these results, the numbers fall in line with what would be
+the logical expectation under these circumstances. The 2013-2014
+structure moves the balance of teams over towards the East with 16 teams
+while the West sits at 14 teams. With more teams in the East, it makes
+sense that the East would have more teams with higher number of total
+points from the season. To look at these results in a different light,
+it is key to translate these numbers to the point scoring system of the
+NHL. On average, the difference between the conference gaps of the two
+division alignments (2012-2013,2013-2014) comes out to be 2.32 in favor
+of the 2013-2014 alignment/structure. Under the 2017-2018 analyis our
+difference in conference gap between alignments comes out to be 1.053
+(1.091-0.038). This result makes sense since the conferences are still
+unbalanced with the East having more teams than the West (16 and 15
+respectively) but less unbalanced than 2013-2014 (16 and 14
+respectively) The conference gap has been calculated as:
 
@@
 
-The mean conference gap, in terms of total points, means that the East’s 8th seeded team has on average 2.30 more points than the West’s 8th seeded team for 2013-2014, similarly we expect the 8th seed in the Eastern conference to have 1.091 more points than their Western conference counterpart under the 2017-2018 structure.
-In the NHL, a team earns 2 points for a win, 1 point for an overtime loss, and 0 points for a loss.
-With this point system,under the 2013-2014 alignment the East’s 8th seeded team on average needs to win one more game during the season or exchange two losses for two losses in overtime.
-This favors Western conference teams in the way they have to earn less total points throughout the season to make the playoffs than Eastern conference teams.
-This same bias towards Western conference teams to make the playoffs is present under the 2017-2018 bias but the practical effects of a one point difference is not even worth a single win in an 82 game season, so although this effect is present it's not as drastic.
+The mean conference gap, in terms of total points, means that the East's
+8th seeded team has on average 2.30 more points than the West's 8th
+seeded team for 2013-2014, similarly we expect the 8th seed in the
+Eastern conference to have 1.091 more points than their Western
+conference counterpart under the 2017-2018 structure. In the NHL, a team
+earns 2 points for a win, 1 point for an overtime loss, and 0 points for
+a loss. With this point system,under the 2013-2014 alignment the East's
+8th seeded team on average needs to win one more game during the season
+or exchange two losses for two losses in overtime. This favors Western
+conference teams in the way they have to earn less total points
+throughout the season to make the playoffs than Eastern conference
+teams. This same bias towards Western conference teams to make the
+playoffs is present under the 2017-2018 bias but the practical effects
+of a one point difference is not even worth a single win in an 82 game
+season, so although this effect is present it's not as drastic.
 
-Not only were the 8th seeds from each conference calculated but each seed for the whole NHL.
-These box and whisker plots show a wide range of information for the total points for each seed across the 10,000 simulated seasons for both the 2013-2014 and 2017-2018 alignemnts.
-The endpoints of the whiskers represent the outlier point totals earned by each seed.
-The sides of the box represent the 2.5% and 97.5% quantiles.
-The line within the box represents the mean.
-The structure of this plot makes sense with the curve of less points the lower seed and more points the higher seed the team achieves.
-The higher the number on the x axis the lower seed of the team; 30 being the worst team in the league (31 for 2017-2018 simulation) and 1 being the best team in the league.
-Interesting feautures of these plots are a tight spread of simulated points for teams seeded around the middle of the league.
-This makes sense given that these point values are probably closer to the mean point total for a season and thus more likely under the model we defined.
-Another note is that the overall range of point totals is greater in the 2013-2014 simulation than it is under the 2017-2018 simulation.
-This has been compared with real NHL data and the shape lines up in a similar fashion.
+Not only were the 8th seeds from each conference calculated but each
+seed for the whole NHL. These box and whisker plots show a wide range of
+information for the total points for each seed across the 10,000
+simulated seasons for both the 2013-2014 and 2017-2018 alignemnts. The
+endpoints of the whiskers represent the outlier point totals earned by
+each seed. The sides of the box represent the 2.5\% and 97.5\%
+quantiles. The line within the box represents the mean. The structure of
+this plot makes sense with the curve of less points the lower seed and
+more points the higher seed the team achieves. The higher the number on
+the x axis the lower seed of the team; 30 being the worst team in the
+league (31 for 2017-2018 simulation) and 1 being the best team in the
+league. Interesting feautures of these plots are a tight spread of
+simulated points for teams seeded around the middle of the league. This
+makes sense given that these point values are probably closer to the
+mean point total for a season and thus more likely under the model we
+defined. Another note is that the overall range of point totals is
+greater in the 2013-2014 simulation than it is under the 2017-2018
+simulation. This has been compared with real NHL data and the shape
+lines up in a similar fashion.
 
-## 2013-2014 Points by Seed
-![](Images/SeedPlace.png){width=80%}
+\hypertarget{points-by-seed}{%
+\subsection{2013-2014 Points by Seed}\label{points-by-seed}}
 
-## 2017-2018 Points by Seed
-![](Images/seedplace20172018.png){width=80%}
+\includegraphics[width=0.8\textwidth,height=\textheight]{Images/SeedPlace.png}
 
-# Conclusion
+\hypertarget{points-by-seed-1}{%
+\subsection{2017-2018 Points by Seed}\label{points-by-seed-1}}
 
-Since the change to the conference structure of the NHL to be unbalanced, there have been questions on whether these structures are giving an unfair advantage for one of the conferences: more specifically the Western Conference.
-With the West having two less teams under the 2013-2014 alignment, it was inferred that each of those teams would have a better chance of making the playoffs.
-They wouldn’t need to earn as many points to secure a playoff spot as teams in the Eastern Conference.
-After running 10,000 Monte Carlo simulations of the 2013-2014 NHL season, the results support the case that the Western Conference has a noticeable advantage over the Eastern Conference.
-The 8th seed in the east on average has an end of season point total of 2.3 (1.091 for 2017-2018) points higher than the 8th seed in the west.
-This can cause problems because a statistically worse team can make the playoffs in the Western Conference while a better team might miss out in the Eastern Conference.
-This is support of the idea that these rules introduce a bias agaisnt teams in the Eastern conference which goes agaisnt the league being "fair" to all teams.
+\includegraphics[width=0.8\textwidth,height=\textheight]{Images/seedplace20172018.png}
 
-# References
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
-- Pettigrew, Stephen.
-“How the West Will Be Won: Using Monte Carlo Simulations to Estimate the Effects of NHL Realignment.” Journal of Quantitative Analysis in Sports, vol.
-10, no.
-3, 1 Jan.
-2014, 10.1515/jqas-2013-0125.
-Accessed 10 Dec.
-2020.
+Since the change to the conference structure of the NHL to be
+unbalanced, there have been questions on whether these structures are
+giving an unfair advantage for one of the conferences: more specifically
+the Western Conference. With the West having two less teams under the
+2013-2014 alignment, it was inferred that each of those teams would have
+a better chance of making the playoffs. They wouldn't need to earn as
+many points to secure a playoff spot as teams in the Eastern Conference.
+After running 10,000 Monte Carlo simulations of the 2013-2014 NHL
+season, the results support the case that the Western Conference has a
+noticeable advantage over the Eastern Conference. The 8th seed in the
+east on average has an end of season point total of 2.3 (1.091 for
+2017-2018) points higher than the 8th seed in the west. This can cause
+problems because a statistically worse team can make the playoffs in the
+Western Conference while a better team might miss out in the Eastern
+Conference. This is support of the idea that these rules introduce a
+bias agaisnt teams in the Eastern conference which goes agaisnt the
+league being ``fair'' to all teams.
 
-- History of Organizational Changes in the NHL. 26 Nov. 2020, en.wikipedia.org/wiki/History_of_organizational_changes_in_the_NHL. Accessed 10 Dec. 2020.
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\begin{itemize}
+\item
+  Pettigrew, Stephen. ``How the West Will Be Won: Using Monte Carlo
+  Simulations to Estimate the Effects of NHL Realignment.'' Journal of
+  Quantitative Analysis in Sports, vol.~10, no. 3, 1 Jan.~2014,
+  10.1515/jqas-2013-0125. Accessed 10 Dec.~2020.
+\item
+  History of Organizational Changes in the NHL. 26 Nov.~2020,
+  en.wikipedia.org/wiki/History\_of\_organizational\_changes\_in\_the\_NHL.
+  Accessed 10 Dec.~2020.
+\end{itemize}
+
+\end{document}
